---
title: "R Notebook"
output: html_notebook
---

#set up


```{r}
library(lme4) #mixed effect linear models

library(emmeans)

library(tidyverse)

library(readr)

library(nlme) #non linear models

library(ggpmisc)

library(multcomp)

library(cowplot)
```

#read CSV


```{r}

allometry_total <- read_csv("allometry/allometry_total.csv")%>%
 mutate(
   landform = as.factor(landform),

    rainfall = factor(rainfall, levels = c( "low", "moderate", "high")))

summary(allometry_total)

```


# construct data set 





```{r}

#calcualte basal area
allo_basal <- allometry_total %>%
  gather("d10", "diamiter", starts_with ("d10"))%>%
  mutate(basal_area = (diamiter*0.5)^2*3.14159265359)%>%
  arrange(tree)%>%
  filter(!is.na(diamiter))

  
 
 

```


```{r}

summary(allo_basal)

```


# basal area vs height

```{r}

allo_basal%>%
  ggplot( aes( basal_area, tree_height, colour = rainfall))+
  geom_point(aes(shape=factor(landform)))+
  geom_smooth(method = "nls",
              formula = y ~ c*x^z,
             se = FALSE,
             method.args = list(start = list(c=125, z=0.383739)))

allo_basal%>%
  ggplot( aes( basal_area, tree_height, colour = rainfall))+
  geom_point()+
  geom_smooth(method = "lm",
              se = F)+
  scale_y_log10()+
  scale_x_log10()

allo_basal%>%
  ggplot( aes( rainfall, tree_height, colour = landform))+
  geom_boxplot()

```

## check slope and intercep for basal area vs height relationship.

```{r}

g.allo.basal <- filter(allometry_biomass, rainfall == "high")

g.allo.model <- lm(log(tree_height) ~ log(basal_area), data = g.allo.basal)

summary(g.allo.model)

```


```{r}
n.s.allo.basal <- filter(allometry_biomass, rainfall != "high")

n.s.allo.lm <- lm(log(tree_height) ~ log(basal_area), data = n.s.allo.basal)

summary(n.s.allo.lm)
```


***very similar intercept and slope to basal area to height relationship found in plot level analysis***

## basal area to hight anova 

```{r}
allo.h.b.aov <- aov(log(tree_height) ~ log(basal_area) + rainfall + landform %in% rainfall, data = allometry_biomass)
summary(allo.h.b.aov)
TukeyHSD(allo.h.b.aov, which =  "rainfall")
```

```{r}
qqnorm(residuals(allo.h.b.aov), pch = 1, frame = FALSE)
qqline(residuals(allo.h.b.aov), col = "steelblue", lwd = 2)
```


```{r}
scatter.smooth(residuals(allo.h.b.aov) ~ fitted(allo.h.b.aov))
```

***confirms differences in basal area vs height relationship observed in plot level data*** 


#  height vs canopy ANCOVA


```{r}
canopyvsheight <- allometry_biomass%>%
  ggplot( aes( tree_height, canopy_volume, colour = rainfall))+
  geom_point(aes(shape = landform))+
  geom_smooth(method = "nls",
              formula = y ~b +x^m,
              se = F,
              method.args = list(start = list(m=3, b=0)))+
  ggtitle("b")+
  xlab("height (m)")+
  ylab(bquote("canopy volume  " (cm^3)))+
  theme_cowplot()+
  theme(legend.position = "none",
        axis.title=element_text(size=12))


print(canopyvsheight)

#ggsave("canopyvsheight.png")
```



```{r}
canopy_logs <- lm( log(canopy_volume) ~ log(tree_height), data = allometry_biomass )


summary(canopy_logs)

exp(0.04307)

```






```{r}


allometry_biomass%>%
  ggplot( aes(tree_height, canopy_volume, colour = rainfall))+
  geom_point(aes(shape=factor(landform)))+
  scale_x_log10()+
  scale_y_log10()+
  geom_smooth(method = "lm")



allometry_biomass%>%
  ggplot( aes(rainfall, log(canopy_volume), colour = landform))+
  geom_boxplot( notch = F)


```




```{r}

cv.height.int <- lm(log(canopy_volume) ~ log(tree_height) * rainfall  *  landform , data = allometry_biomass)

summary(cv.height.int)


```


```{r}

cv.height.norm <- lm(log(canopy_volume) ~ log(tree_height) + rainfall + landform , data = allometry_biomass)


AIC(cv.height.int,cv.height.norm)

anova(cv.height.int,cv.height.norm,test = "LRT")


```
***given constant slop across sites indicated by the lack of effect of an interaction on model performance, we can run an check for differences in intercept between sites***





```{r}

pairs(emmeans(cv.height.norm, ~ rainfall))

```

***difference between high and other sites. no difference in slope***



## height vs canopy volume coeficients 





## check assumptions 




```{r}
qqnorm(residuals(cv.height.norm), pch = 1, frame = FALSE)
qqline(residuals(cv.height.norm), col = "steelblue", lwd = 2)
```

```{r}
scatter.smooth(residuals(cv.height.norm) ~ fitted(cv.height.norm))
```


## tree height vs canopy coefficients

### gunda booka 


```{r}
gunda_allo <- filter(allometry_biomass, rainfall == "high")


gunda_logs <- lm( log(canopy_volume) ~ log(tree_height), data = gunda_allo )


summary(gunda_logs)

exp(-0.9756)

```
*y = 0.42 + x^2.9997
 
 
 
 
### nocoleche and sturt
 
  
```{r}
low_allo <- filter(allometry_biomass, rainfall != "high")


noco_allo <- filter(low_allo, rainfall == "moderate")

sturt_allo <- filter(low_allo, rainfall == "low")

```



```{r}

noco_logs <- lm(log(canopy_volume) ~ log(tree_height), data = noco_allo )

summary(noco_logs)
exp(-0.08908)

```
***y = 85 *x^3.6644***




```{r}

sturt_logs <- lm(log(canopy_volume) ~ log(tree_height), data = sturt_allo )

summary(sturt_logs)
exp(0.09407)

```



## plot curves of best fit


### gunda 

***use equation from log plots above***
```{r}
gunda_allo%>%
  ggplot(aes(tree_height, canopy_volume))+
  geom_point()+
  theme_bw()+
  geom_smooth(method = "nls",
              formula = y ~ b + x^m,
             se = FALSE,
             method.args = list(start = list(b= 0, m = 2.9997 )))
```     
 



 


# canopy volume vs basal area  

 geom_smooth(method = "nls",
              formula = y ~b +x^m,
              se = F,
              method.args = list(start = list(m=3, b=0)))+

```{r}
allometry_biomass%>%
  ggplot( aes( basal_area, canopy_volume, colour = rainfall))+
  geom_point(aes(shape=factor(landform)))+
   geom_smooth(method = "nls",
              formula = y ~ b + x^m,
              se = FALSE,
              method.args = list(start = list(b= 0.4660558, m = 1.13)))


```


```{r}

cv.ba.params<- lm(log(canopy_volume)~log(basal_area), data = allometry_biomass)

summary(cv.ba.params)

exp(-0.76345)
```




```{r}

basalvscanopy <- allometry_biomass%>%
  ggplot( aes( basal_area, canopy_volume, colour = rainfall))+
  geom_point(aes(shape = landform))+
     geom_smooth(method = "nls",
              formula = y ~ b + x^m,
              se = FALSE,
              method.args = list(start = list(b= 0.4660558, m = 1.13)))+
  ylab(bquote("canopy volume " (cm^3)))+
  xlab(bquote("basal area " (cm^2)))+
  ggtitle("c")+
  theme_cowplot()+
  theme(axis.title=element_text(size=12,face="bold"))+
  theme(legend.position = "none")
  


 print(basalvscanopy)

#ggsave("basalvscanopy.png")

```

***suggests as basal area increases canopy volume increases exponentially. investement in basal area returns considerable increase in canopy volume.***





```{r}
lm.canopy_area <- lm(log(canopy_volume) ~ log(basal_area), data = allometry_biomass)

summary(lm.canopy_area)
```







```{r}

cv.basal.inter <- lm(log(canopy_volume) ~ log(basal_area) + rainfall + landform, data = allometry_biomass)

summary(cv.basal.inter)


```

```{r}
cv.basal.norm <- lm(log(canopy_volume) ~ log(basal_area)  + rainfall + landform, data = allometry_biomass)

AIC(cv.basal.inter,cv.basal.norm)

anova(cv.basal.inter,cv.basal.norm,test = "LRT")

```

```{r}

pairs(emmeans(cv.basal.norm, ~ rainfall))

```



## check assumptions

```{r}
qqnorm(residuals(cv.basal.norm), pch = 1, frame = FALSE)
qqline(residuals(cv.basal.norm), col = "steelblue", lwd = 2)
```

```{r}
scatter.smooth(residuals(cv.basal.norm) ~ fitted(cv.basal.norm))
```





***y= 0 + x^1.13520**




```{r}
allometry_biomass%>%
  ggplot(aes(basal_area, canopy_volume))+
  geom_point()+
  theme_bw()+
  scale_y_log10()+
  scale_x_log10()+
  geom_smooth(method = "lm",
              se = T)
```


```{r}
gunda_cv.basal<- lm(log(canopy_volume) ~ log(tree_height), data = gunda_allo)

summary(gunda_cv.basal)
```

```{r}
noco_cv.basal<- lm(log(canopy_volume) ~ log(tree_height), data = noco_allo)

summary(noco_cv.basal)
```















