---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# plots


# set up

```{r}

library(lme4) #mixed effect linear models

library(lmerTest)

library(emmeans)

library(tidyverse)

library(readr)

library(nlme)

library(ggpmisc)

library(cowplot)

library(car)

library(forcats)

library(mgcv)

library(glmmLasso)

```

```{r}
total_plots <- read_csv("plot_data/total_plots.csv")%>%
  mutate(
   landform = as.factor(landform),

    plots = as.factor(plots),
   
   subplot = as.factor(subplot),

    rainfall = factor(rainfall, levels = c("low", "moderate" , "high")))

summary(total_plots)
```
# build data set

# calcuale basal area and summed biomass


```{r}

 #calculate basal area   
  basal_area <- total_plots %>%
  gather("d_10", "diamiter", starts_with ("d_"))%>%
  mutate(basal_area = (diamiter*0.5)^2*3.14159265359)%>%
  arrange(tree)%>%
  filter(!is.na(diamiter))
```

#d2h model


M(kg) = 34.4703 - 8.0671 DBH + 0.6589 DBH^2 (cm)

```{r}
  
 # using the accacia allometric equation from feller 1980 we will calculate biomass for each tree based on a "stem by stem" analysis.
# see "Review of Allometric Relationships for Estimating Woody Biomass for New South Wales, the Australian Capital Territory, Victoria, Tasmania and South Australia" (page 36)
    D2H_model <- basal_area %>%
   group_by(tree, height, landform, plots, rainfall, subplot, sub_plot_area, subplot_density, plot_density)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
      mutate(diamiter = (sqrt(basal_area/3.14159265359))*2,
             dbh = (diamiter*0.798)-0.577,
             stem_log_biomass = log(1.9+ 424.9 * (dbh/100)^2 * (height/100)),
             stem_biomass = exp(stem_log_biomass),
             branch_log_biomass = log(208) + log((dbh/100)^2 * (height/100)),
             branch_biomass = exp(branch_log_biomass),
             leave_log_biomass = log(8) + log((dbh/100)^2 * (height/100)),
             leave_biomass = exp(leave_log_biomass))%>%
      mutate(
        biomass = stem_biomass + branch_biomass + leave_biomass,
        log_biomass = log(biomass))

```
    

#d2h plot level biomass

```{r}

    
  plot_D2h <- D2H_model%>%
  group_by( subplot, landform, sub_plot_area, plots, rainfall, subplot_density, plot_density)%>%
    summarise(
      ave.biomass = mean(biomass),
      diameter = mean(diamiter),
      tot.biomass = sum(biomass),
      stem_area = mean(3.1415/4*diamiter^2),
    biomass.m2 = tot.biomass/sub_plot_area)%>%
  ungroup()%>%
  mutate(
    logbiomass.m2 = log(biomass.m2),
    log.sub.dens = log(subplot_density),
    plots = as.numeric(plots))

summary(plot_D2h)

```






```{r}
names(plot_D2h)
```


```{r}
stand_metrics <- plot_D2h %>% 
  group_by( subplot, landform, sub_plot_area, plots, rainfall, subplot_density, plot_density)%>%
  summarise(ave.biomass = mean(ave.biomass),
            logbiomass.m2 = mean(logbiomass.m2),
            log.sub.dens = mean(log.sub.dens),
            biomass.m2 = mean(biomass.m2))
  


summary(stand_metrics)
```




# combinded diameter only model 

```{r}
#combind basal area for multi stemmed trees for a "combined stem analysis" 
 ind_biomass <- basal_area%>%
  group_by(tree, height, landform, plots, rainfall, subplot, sub_plot_area, subplot_density, plot_density)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
  mutate(
     #back tranform total_basal_area to diamiter 
    diamiter = case_when( rainfall == "low" ~ ((sqrt(basal_area/3.14159265359))*2),
                          rainfall == "moderate" ~ ((sqrt(basal_area/3.14159265359))*2),
                          rainfall == "high" ~ ((sqrt(basal_area/3.14159265359))*2)*0.798)-0.577, # converting d10 to d130
     # Equations taken from -> Paul et al, "testing the generality of above ground ind_biomass allometry across plant functional types at the continent scale" -> Fmulti "ln(AGB) = 2.474ln(D10)-2.757" used for lower rainfall sites (page 9)
    logAGB = case_when(rainfall == 'low' ~ 2.474*log(diamiter)-2.757,
                      rainfall == 'moderate' ~ 2.474*log(diamiter)-2.757,
                      rainfall == 'high' ~ 2.220*log(diamiter)-1.693),
    # correctional factors taken form relevant equations in respective papers. CF for high rainfall taken form average of Acacia trees paul and roxburgh 2013.
    AGB = case_when( rainfall == 'low' ~ exp(logAGB)*1.0787,
                     rainfall == 'moderate' ~ exp(logAGB)*1.0787,
                     rainfall == 'high' ~ exp(logAGB)*1.0436))
 
    
 #plot level biomass
plot_biomass <- ind_biomass%>%
  aggregate(AGB ~ subplot + landform + sub_plot_area + plots + rainfall + subplot_density + plot_density, sum)%>%
  mutate(
    biomass.m2 = AGB/sub_plot_area,
    logbiomass.m2 = log(biomass.m2),
    log.sub.dens = log(subplot_density))

```



```{r}
#combind basal area for multi stemmed trees for a "combined stem analysis" 
 Tree_model <- basal_area%>%
  group_by(tree, height, landform, plots, rainfall, subplot, sub_plot_area, subplot_density, plot_density)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
  mutate(
     #back tranform total_basal_area to diamiter 
    diamiter =  (((sqrt(basal_area/3.14159265359))*2)*0.798)-0.577,
    logAGB = 2.220*log(diamiter)-1.693,
    AGB = exp(logAGB)*1.0436)
    
 #plot level biomass
tree_plot <- diamiter_model%>%
  aggregate(AGB ~ subplot + landform + sub_plot_area + plots + rainfall + subplot_density + plot_density, sum)%>%
  mutate(
    biomass.m2 = AGB/sub_plot_area,
    logbiomass.m2 = log(biomass.m2),
    log.sub.dens = log(subplot_density))

```

#Paul Shrub

```{r}

#individual level

paul_shrub <- basal_area%>%
  group_by(tree, height, landform, plots, rainfall, subplot, sub_plot_area, subplot_density, plot_density)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
  mutate( 
    diamiter = ((sqrt(basal_area/3.14159265359))*2),
    p.logAGB = 2.474*log(diamiter)-2.757,
    p.AGB = exp(p.logAGB)*1.0787)


#plot level 
    
p.shrub_plots <- paul_shrub%>%
  aggregate(p.AGB ~ subplot + landform + sub_plot_area + plots + rainfall + subplot_density + plot_density, sum)%>%
  mutate(
    biomass.m2 = p.AGB/sub_plot_area,
    logbiomass.m2 = log(biomass.m2),
    log.sub.dens = log(subplot_density))


```

```{r}
summary(diamiter_model)
```



```{r}
 summary(D2H_model)
```



```{r}
 summary(ind_biomass)
```




```{r}
isocline <- ind_biomass %>%
  group_by(subplot,  landform, sub_plot_area, plots , rainfall)%>%
  summarise(
    
    basal_area = sum(basal_area),
    subplot.n = n(),
    subplot_density = subplot.n / sub_plot_area[1],
    ave.AGB = mean(AGB),
    ) %>% ungroup()


isocline$rainfall <-  factor(isocline$rainfall, 
                             levels = c("low", "moderate" , "high"))


summary(isocline)

  
```


```{r}

ind_biomass%>%
  mutate(basal_area = as.numeric(basal_area))

summary(ind_biomass)

```



```{r}

summary(plot_biomass)

```



# filtering
```{r}
f.ind_biomass <- filter(ind_biomass, diamiter > 2)

f.D2H_model <- filter(D2H_model, diamiter > 2)

f.diamiter_model <- filter(diamiter_model, diamiter >2)

```



```{r}
d2h.h.rain. <- filter(f.D2H_model, rainfall == "high")
diam.h.rainf <- filter(ind_biomass, rainfall == "high")
```




```{r}

#plot damiter measurements 

#diamiter
diam.h.rainf%>%
ggplot( aes(x= AGB)) +
  geom_histogram()+
  ggtitle("diameter only")
  

d2h.h.rain.%>%
  ggplot( aes(x = biomass))+
  geom_histogram()+
  ggtitle("d2h model")
  
```





```{r}

#plot biomass measurements 

#biomass
f.ind_biomass%>%
ggplot(aes(x=AGB, colour = factor(rainfall))) +
  geom_histogram( fill = "#69b3a2",
                  alpha = 0.09,
                  bins = 40)+
  ggtitle("summed basal")+
  scale_x_log10()+
  theme_cowplot()+
  theme(plot.title = element_text(size = 15))

#D2h biomass model

f.D2H_model %>%
  ggplot( aes( x = biomass, colour = factor (rainfall)))+
   geom_histogram( fill = "#69b3a2",
                  alpha = 0.09,
                  bins = 40)+
  scale_x_log10()+
  ggtitle("stem by stem ")+
  theme_cowplot()+
  theme(plot.title = element_text(size = 15))
```


```{r}

#plot the ind_biomass dist

#multistem

ind_biomass%>%
ggplot( aes(x=AGB, colour = factor(rainfall))) +
  geom_histogram( fill = "#69b3a2",
                  alpha = 0.09,
                  bins = 40)+
  scale_x_log10()+
  ggtitle("log scale ind_biomass")+
  theme_classic()+
  theme(plot.title = element_text(size = 15))

ind_biomass%>%
  ggplot( aes(x=AGB, colour = factor(rainfall)))+
            geom_histogram(bins = 40)+
            theme_grey()+
            ggtitle("Normal Scale")+
     
         theme(plot.title = element_text(size = 15))
```


# basal area and height 



```{r}

fb.d2H <- D2H_model%>%
  filter(basal_area < 1450)
summary(fb.d2H)

```

```{r}
appendix.structure <- fb.d2H%>%
  ggplot(aes(diamiter, height, colour = rainfall))+
  geom_point(aes(shape = landform))+
geom_smooth(method = "nls",
              formula = y ~ c*x^z,
             se = FALSE,
             method.args = list(start = list(c=125, z=0.383739)))+
  ylab("height (cm)")+
  xlab(bquote("diameter " (cm)))+
  #ggtitle("a")+
  theme_cowplot()+
  theme(axis.title=element_text(size=12))
  

print(appendix.structure)

#ggsave("appendix.structure.png")

```


```{r}
appendix.diam_height <- lmer(log(height) ~ log(diamiter) * rainfall + landform + (1|plots/subplot), data = f.D2H_model, REML = T)

summary(appendix.diam_height)

```



```{r}
qqnorm(residuals(appendix.diam_height), pch = 1, frame = FALSE)
qqline(residuals(appendix.diam_height), col = "steelblue", lwd = 2)

```

```{r}
scatter.smooth(residuals(appendix.diam_height) ~ fitted(appendix.diam_height))
```




```{r}
# null vs fit

appendix.diam_height.anova <- lmer(log(height) ~ log(diamiter) * rainfall + landform +  (1|plots/subplot) , data = f.D2H_model, REML = F)
appendix.diam_height.null <- lmer(log(height) ~ log(diamiter) + landform + (1|plots/subplot), data = f.D2H_model, REML = F)

anova(appendix.diam_height.null,appendix.diam_height.anova,test = "LRT")

```




```{r}
#differences in slope/ineraction effects

anova(appendix.diam_height.anova)


```


```{r}
#difference in intercept between high and other two sites

 pairs(emmeans(appendix.diam_height, ~ rainfall*diamiter))
```


```{r}
#difference in intercept between high and other two sites

 pairs(emmeans(appendix.diam_height.anova, ~ rainfall))
```



# standardized biomass esitamtes 



```{r}
exp(2.757)

log(15.75251)


```


```{r}


#model calibration 

#general model is calibrated on "species-specific" fitted model. We should calibrate this at the low rainfall site when hight is the lowest, and will have the least effect on the biomass estimates. 

# for species specific model.............AGB = a1 * D^b (b = 2.474), (a1 = 15.75251)

# for the general model .................AGB = a2 * (D^2*H)

# therefore .............................a1 + D^b = a2 * (D^2*H)

# for a2 ................................a2 = a1 * D^b /(D^2*H)

# to calibrate this on the low rainfall site, apply this transformation "a1 + D^b/((D^2*H)^c/100)" at the low rainfall site. The average will produce a1.Given, a1, a2 can be calibrated.THis will be the intercepts for general model.


 calibration_DF <- basal_area %>%
   group_by(tree, height, landform, plots, rainfall, subplot, sub_plot_area, subplot_density, plot_density)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
      mutate(diamiter = (sqrt(basal_area/3.14159265359))*2, #calculated diameter form summed basal area
             a2 = (-1.9 + 15.75251 * diamiter^2.474)/(diamiter^2 * height)) # transformation taken form above equation 

summary(calibration_DF)

             

```








```{r}
# determine the mean value of a2 at the low rainfall site 

l.rain. <- filter(calibration_DF, rainfall != "high")

low.rain <- filter(l.rain., rainfall != "moderate")

summary(low.rain) # mean a2 = 0.12677
```

~ 2.474*log(diamiter)-2.757


```{r}
# new equation is AGB = 0.12709   * D^2 * (H)
# D and H are in meters in the equation but centimeters in data set

cal_gen <- basal_area %>%
   group_by(tree, height, landform, plots, rainfall, subplot, sub_plot_area, subplot_density, plot_density)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
      mutate(diamiter = (sqrt(basal_area/3.14159265359))*2,
             dbh = (diamiter*0.798)-0.577,     #scaling factor to turn d10 to dbh 
             g.logAGB = log(1.9 + 0.12677 * ((dbh)^2 * (height))/100),
             g.AGB = exp(g.logAGB))

summary(cal_gen)

```


```{r}

cal_gen_plots <- cal_gen%>%
    group_by( subplot, landform, sub_plot_area, plots, rainfall, subplot_density, plot_density)%>%
    summarise(
      ave.biomass = mean(g.AGB),
      diameter = mean(diamiter),
      tot.biomass = sum(g.AGB),
      stem_area = mean(3.1415/4*diamiter^2),
    biomass.m2 = tot.biomass/sub_plot_area)%>%
  ungroup()%>%
  mutate(
    logbiomass.m2 = log(biomass.m2),
    log.sub.dens = log(subplot_density),
    plots = as.numeric(plots))

```



## filtering
```{r}
f.cal_gen <- filter(cal_gen, diamiter > 2)

f.paul_shrub <- filter(paul_shrub, diamiter > 2)
```


# VISUAL COMPARISON 

```{r}

#comparison of stand level biomass


p.stand<- p.shrub_plots%>%
  ggplot(aes(rainfall, biomass.m2, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab(bquote("biomass "(kg ~ m^-2)))+
  xlab( " Mulga specific stand level ")+
  #ggtitle("tree")+
  theme_cowplot()+
  theme(legend.position = "none")
print(p.stand)


g.stand <- cal_gen_plots%>%
  ggplot(aes(rainfall, biomass.m2, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab(bquote(""))+
  xlab( " General stand level ")+
  #ggtitle("tree")+
  theme_cowplot()+
  theme(legend.position = "none")
print(g.stand)

```



```{r}
# comparison of individual tree bioamss 

shrub_ind_dist <- paul_shrub%>%
  ggplot( aes(p.AGB, fill = rainfall ))+
  geom_histogram()+
  ggtitle("Mulga specific")+
  xlim(c(NA,1000))+
  ylim(c(NA, 1000))+
  guides(fill="none")+
  xlab("Biomass in kg")+
  theme_cowplot()
print(shrub_ind_dist)


tree_ind_dist <- cal_gen%>%
  ggplot( aes(g.AGB, fill = rainfall ))+
  geom_histogram()+
  ggtitle("General equation")+
  xlim(c(NA, 1000))+
  ylim(c(NA, 1000))+
  guides(fill="none")+
  xlab(" Biomass in kg ")+
  theme_cowplot()
print(tree_ind_dist)



```







```{r}
#plot damiter measurements 

#diamiter
diam.h.rainf%>%
ggplot( aes(x= AGB)) +
  geom_histogram()+
  ggtitle("diameter only")
  

d2h.h.rain.%>%
  ggplot( aes(x = biomass))+
  geom_histogram()+
  ggtitle("d2h model")
```






```{r}
paul.h.rain <- filter(f.paul_shrub, rainfall == "high")
gen.h.rainf <- filter(f.cal_gen, rainfall == "high")
```


```{r}
paul.h.rain%>%
    ggplot( aes(p.AGB))+
  geom_histogram()+
  ggtitle("shrub")


gen.h.rainf%>%
  ggplot( aes(g.AGB))+
  geom_histogram()+
  ggtitle("d2h")


```







```{r}

plot_grid(shrub_ind_dist, tree_ind_dist, rel_widths = c(1,1))

ggsave("dist_comparison.png", width = 12, height =5)
  
```


# kolomogrov-smirnov test (compare distributions)

## create vectors 

```{r}
# extract biomass as a vector form the paul and genneral DFs


paul.bio <- f.paul_shrub$p.AGB
gen.bio <- f.cal_gen$g.AGB 


```




## ks.test 

```{r}
ks.test(gen.bio, paul.bio)
```



# Largest 20% comparison


```{r}

# paul 
paul_20 <- f.paul_shrub %>%
  arrange(desc(p.AGB)) %>%       
  mutate(rank = row_number()) %>%
  filter(rank <= n() * 0.2) %>%
  select(-rank)                           

print(paul_20)


```



```{r}
# general 

gen_20 <- f.cal_gen %>%
  arrange(desc(g.AGB)) %>%       
  mutate(rank = row_number()) %>%
  filter(rank <= n() * 0.2) %>%
  select(-rank)                           

print(gen_20)


```

## compare totals 


```{r}

paul_20_total <- sum(paul_20$p.AGB)

gen_20_total <- sum(gen_20$g.AGB)


print(paul_20_total)

print(gen_20_total)

```

## compare percentages 

```{r}
total_paul <- sum(f.paul_shrub$p.AGB)

total_gen <- sum(f.cal_gen$g.AGB)

print(total_paul)

print(total_gen)
```

```{r}
paul_20_percentage <- 55733.43/84314.47

gen_20_percentage <- 90028.48/137332.2

print(paul_20_percentage)

print(gen_20_percentage)


```


## comparative make up 

```{r}
high_paul_20 <-  filter(paul_20, rainfall == "high")

high_gen_20 <- filter(gen_20, rainfall == "high")


print(high_paul_20)

print(high_gen_20)
```


```{r}
69/354

128/354
```


```{r}

plot_grid(p.stand, g.stand, rel_widths = c(1.1,1))

ggsave("standlevelcomparison.png", width = 12, height =4)
  
```





