

# STATs for Mulga trait shift paper 



# set up

```{r}

library(lme4) #mixed effect linear models

library(lmerTest)

library(emmeans)

library(tidyverse)

library(readr)

library(nlme)

library(ggpmisc)

library(cowplot)

library(car)

library(forcats)

library(mgcv)

library(glmmLasso)

library (glmmTMB)

library(sp)

library(gstat)

library(ape)

```

#INDIVIUDAL TREE LEVEL METRICS

## build data sets

### import/build data set



```{r}
total_plots <- read_csv("plot_data/total_plots.csv")%>%
  mutate(
   landform = as.factor(landform),

    plots = as.factor(plots),
   
   subplot = as.factor(subplot),

    rainfall = factor(rainfall, levels = c("low", "moderate" , "high")))

summary(total_plots)
```



### calcuale basal area and summed biomass



```{r}

 #calculate basal area   
  basal_area <- total_plots %>%
  gather("d_10", "diamiter", starts_with ("d_"))%>%
  mutate(basal_area = (diamiter*0.5)^2*3.14159265359)%>%
  arrange(tree)%>%
  filter(!is.na(diamiter))
```

### biomass estimates for individual trees


```{r}
  
 # using the accacia allometric equation from feller 1980 we will calculate biomass for each tree based on a "stem by stem" analysis.
# see "Review of Allometric Relationships for Estimating Woody Biomass for New South Wales, the Australian Capital Territory, Victoria, Tasmania and South Australia" (page 36)
    D2H_model <- basal_area %>%
   group_by(tree, height, landform, plots, rainfall, subplot, sub_plot_area, subplot_density, plot_density)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
      mutate(diamiter = (sqrt(basal_area/3.14159265359))*2,
             dbh = (diamiter*0.798)-0.577,# tapering equation from Paul et al, "testing the generality of above ground ind_biomass allometry across plant functional types at the continent scale"
             stem_log_biomass = log(1.9+ 424.9 *(dbh/100)^2 * (height/100)), #ensure units of measurement are converted to cm
             stem_biomass = exp(stem_log_biomass),
             branch_log_biomass = log(208) + log((dbh/100)^2 * (height/100)),
             branch_biomass = exp(branch_log_biomass),
             leave_log_biomass = log(8) + log((dbh/100)^2 * (height/100)),
             leave_biomass = exp(leave_log_biomass))%>%
      mutate(
        biomass = stem_biomass + branch_biomass + leave_biomass,
        log_biomass = log(biomass))

```




### filtered 

```{r}
#filter out smaller trees due to allometeric differences in juveniles

f.D2H_model <- filter(D2H_model, diamiter > 2)


```



### allometric data with canopy volume

```{r}

allometry_total <- read_csv("allometry/allometry_total.csv")%>%
 mutate(
   landform = as.factor(landform),

    rainfall = factor(rainfall, levels = c( "low", "moderate", "high")))

summary(allometry_total)

```



### calculate basal area and canopy volume



```{r}

#calcualte basal area
allo_basal <- allometry_total %>%
  gather("d10", "diamiter", starts_with ("d10"))%>%
  mutate(basal_area = (diamiter*0.5)^2*3.14159265359)%>%
  arrange(tree)%>%
  filter(!is.na(diamiter))


```





```{r}

canopy_volume <- allo_basal%>%
  group_by(tree, canopy_1, canopy_2, `95_Canopy`, tree_height, landform, rainfall)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
  mutate(canopy_height = ((tree_height - `95_Canopy`)/2),  # divide by 2 to get the semi-axis
         canopy_length = canopy_1/2,
         canopy_bredth = canopy_2/2,
         canopy_volume = (3/4 * 3.141593 * canopy_height *canopy_bredth * canopy_length)/1000000,# volume of an ellipsoid. Divide by 1,000,000 to get m^3
         log.tree.H = log(tree_height),
         log.canopy.V = log(canopy_volume),
         log.basal_area = log(basal_area))%>%
  filter(!is.na(canopy_volume))%>%
  filter(!is.na(log.canopy.V))

  

```


```{r}

summary(allo_basal)

```




## Individual tree level biomass (LMEM)

```{r}

final.ind.tree.fit <- lmer(log_biomass ~  rainfall  + (1|landform/plots/subplot), data = f.D2H_model, REML = T)

summary(final.ind.tree.fit)

 
```
### assumptions

```{r}
qqnorm(residuals(final.ind.tree.fit), pch = 1, frame = FALSE)
qqline(residuals(final.ind.tree.fit), col = "steelblue", lwd = 2)

```


```{r}
scatter.smooth(residuals(final.ind.tree.fit) ~ fitted(final.ind.tree.fit))
```


```{r}
# Extract the residuals from the model
fin.tree.residuals <- residuals(final.ind.tree.fit)

# Perform the ACF analysis
 acf(fin.tree.residuals)
```



### dredging 

```{r}
# null vs fit

final.ind.tree.fit <- lmer(log_biomass ~  rainfall + (1|landform/plots/subplot), data = f.D2H_model, REML = F)
final.ind.tree.null <- lmer(log_biomass ~ (1|landform/plots/subplot), data = f.D2H_model, REML = F)

anova(final.ind.tree.fit,final.ind.tree.null,test = "LRT")

```

### Hypothesis testing for ranomd effects

```{r}
#test if we need landform given we have plots as a random effect

nBoot <- 1000
lrStat <- rep(NA, nBoot)
final.ind.tree.null <- lmer(log_biomass ~  rainfall + (1|plots/subplot) , data = f.D2H_model, REML = F) #null model
final.ind.tree.fit <- lmer(log_biomass ~  rainfall + (1|landform/plots/subplot) , data = f.D2H_model, REML = F) # full model
lrObs <- 2 * logLik(final.ind.tree.fit) - 2 * logLik(final.ind.tree.null) # observed test stat
for (iBoot in 1:nBoot)
{
  f.D2H_model$logBiomassSim <- unlist(simulate(final.ind.tree.null)) # resampled data
  bNull <- lmer(logBiomassSim ~ rainfall + (1 | plots/subplot), data = f.D2H_model, REML = F) # null model
  bAlt <- lmer(logBiomassSim ~ rainfall + (1 | landform/plots/subplot), data = f.D2H_model, REML = F) # full model
  lrStat[iBoot] <- 2 * logLik(bAlt) - 2 * logLik(bNull) # resampled test stat
}
mean(lrStat > lrObs) # P-value for test of plots effect
```

no evidence to include landform as a random effect

```{r}
#test if we need plots given we have subplots as a random effect

nBoot <- 1000
lrStat <- rep(NA, nBoot)
final.ind.tree.null <- lmer(log_biomass ~  rainfall + (1|subplot) , data = f.D2H_model, REML = F) #null model
final.ind.tree.fit <- lmer(log_biomass ~  rainfall + (1|plots/subplot) , data = f.D2H_model, REML = F) # full model
lrObs <- 2 * logLik(final.ind.tree.fit) - 2 * logLik(final.ind.tree.null) # observed test stat
for (iBoot in 1:nBoot)
{
  f.D2H_model$logBiomassSim <- unlist(simulate(final.ind.tree.null)) # resampled data
  bNull <- lmer(logBiomassSim ~ rainfall + (1 |subplot), data = f.D2H_model, REML = F) # null model
  bAlt <- lmer(logBiomassSim ~ rainfall + (1 |plots/subplot), data = f.D2H_model, REML = F) # full model
  lrStat[iBoot] <- 2 * logLik(bAlt) - 2 * logLik(bNull) # resampled test stat
}
mean(lrStat > lrObs) # P-value for test of plots effect
```


evidence to include plots


## Final individual level biomass model 


```{r}

final.ind.tree.fit <- lmer(log_biomass ~  rainfall  + (1|plots/subplot), data = f.D2H_model, REML = T)

summary(final.ind.tree.fit)

 
```



```{r}
pairs(emmeans(final.ind.tree.fit, ~  rainfall))
```

```{r}

ind.biomass <- f.D2H_model %>%
  ggplot(aes(rainfall, biomass, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab(bquote("biomass"(kg ~ tree^-1)))+
  xlab(" ")+
  ggtitle("d2h")+
  theme_cowplot()+
  theme(legend.position = "none",
        axis.title=element_text(size=12))


print(ind.biomass)
```





## Basal area to height (LMEM)



```{r}
#the interaction term (*) between log(basal_area) and rainfall allows the "height Vs basal areas" relationship to vary between levels of rainfall (or across sites). 

final.height.basal.lm <- lmer(log(height) ~ log(basal_area) * rainfall + (1|landform/plots/subplot), data = f.D2H_model, REML = T)

summary(final.height.basal.lm)

```

### assumptions
```{r}
qqnorm(residuals(final.height.basal.lm), pch = 1, frame = FALSE)
qqline(residuals(final.height.basal.lm), col = "steelblue", lwd = 2)

```


```{r}
scatter.smooth(residuals(final.height.basal.lm) ~ fitted(final.height.basal.lm))
```





```{r}
vif(final.height.basal.lm)
```



```{r}
# Extract the residuals from the model
height.basal.residuals <- residuals(final.height.basal.lm)

# Perform the ACF analysis
 acf(height.basal.residuals)
```


### dredging

```{r}
# null vs fit

final.height.basal.aov <- lmer(log(height) ~ log(basal_area) * rainfall + (1|landform/plots/subplot) , data = f.D2H_model, REML = F)
final.height.null <- lmer(log(height) ~ log(basal_area) + (1|landform/plots/subplot), data = f.D2H_model, REML = F)

anova(final.height.null,final.height.basal.aov,test = "LRT")

```




### hypothesis testing for random effects

```{r}
#test if we need landform given we have plots as a random effect

nBoot <- 1000
lrStat <- rep(NA, nBoot)
final.height.basal.null <- lmer(log(height) ~ log(basal_area) * rainfall + (1|plots/subplot) , data = f.D2H_model, REML = F) #null model
final.height.basal.aov <- lmer(log(height) ~ log(basal_area) * rainfall + (1|landform/plots/subplot) , data = f.D2H_model, REML = F) # full model
lrObs <- 2 * logLik(final.height.basal.aov) - 2 * logLik(final.height.basal.null) # observed test stat
for (iBoot in 1:nBoot)
{
  f.D2H_model$logHeightSim <- unlist(simulate(final.height.basal.null)) # resampled data
  bNull <- lmer(logHeightSim ~ log(basal_area) * rainfall + (1 | plots/subplot), data = f.D2H_model, REML = F) # null model
  bAlt <- lmer(logHeightSim ~ log(basal_area) * rainfall + (1 | landform/plots/subplot), data = f.D2H_model, REML = F) # full model
  lrStat[iBoot] <- 2 * logLik(bAlt) - 2 * logLik(bNull) # resampled test stat
}
mean(lrStat > lrObs) # P-value for test of plots effect
```


```{r}
#test if we need subplots given we have plots as a random effect

nBoot <- 1000
lrStat <- rep(NA, nBoot)
final.height.basal.null <- lmer(log(height) ~ log(basal_area) * rainfall + (1|plots) , data = f.D2H_model, REML = F)
final.height.basal.aov <- lmer(log(height) ~ log(basal_area) * rainfall + (1|plots/subplot) , data = f.D2H_model, REML = F) # full model
lrObs <- 2 * logLik(final.height.basal.aov) - 2 * logLik(final.height.basal.null) # observed test stat
for (iBoot in 1:nBoot)
{
  f.D2H_model$logHeightSim <- unlist(simulate(final.height.basal.null)) # resampled data
  bNull <- lmer(logHeightSim ~ log(basal_area) * rainfall + (1 | plots), data = f.D2H_model, REML = F) # null model
  bAlt <- lmer(logHeightSim ~ log(basal_area) * rainfall + (1 |plots/subplot), data = f.D2H_model, REML = F) # full model
  lrStat[iBoot] <- 2 * logLik(bAlt) - 2 * logLik(bNull) # resampled test stat
}
mean(lrStat > lrObs) # P-value for test of plots effect
```





## Final basal area to height model 

```{r}
final.height.basal.lm <- lmer(log(height) ~ log(basal_area) * rainfall + (1|plots/subplot) , data = f.D2H_model, REML = T)

summary(final.height.basal.lm)
```


```{r}
pairs(emmeans(final.height.basal.lm, ~  rainfall))
```

```{r}
basal.vs.heigh_plots <- f.D2H_model%>%
  ggplot(aes(basal_area, height, colour = rainfall))+
  geom_point(aes(shape = landform))+
geom_smooth(method = "nls",
              formula = y ~ c*x^z,
             se = FALSE,
             method.args = list(start = list(c=125, z=0.383739)))+
  ylab("height (cm)")+
  xlab(bquote("basal area " (cm^2)))+
  xlim(c(NA, 1250))+
  ggtitle("a")+
  theme_cowplot()+
  theme(axis.title=element_text(size=12), legend.position = "none")
  

print(basal.vs.heigh_plots)

```


## Tree height (LMEM)

```{r}
final.tree.height.lm <- lmer(log(height) ~  rainfall + (1|landform/plots/subplot), data = f.D2H_model, REML = T)

summary(final.tree.height.lm)
```


### assumptions

```{r}

qqnorm(residuals(final.tree.height.lm), pch = 1, frame = FALSE)
qqline(residuals(final.tree.height.lm), col = "steelblue", lwd = 2)


```



```{r}
scatter.smooth(residuals(final.tree.height.lm) ~ fitted(final.tree.height.lm))
```


```{r}
# Extract the residuals from the model
residuals <- residuals(final.tree.height.lm)

# Perform the ACF analysis
 acf(residuals)
```




### dredging

```{r}
final.tree.height.lm <- lmer(log(height) ~ rainfall + (1|landform/plots/subplot) , data = f.D2H_model, REML = F)

final.tree.height.null <- lmer(log(height) ~  (1|landform/plots/subplot) , data = f.D2H_model, REML = F)

anova(final.tree.height.null,final.tree.height.lm,test = "LRT")
```


***rainfall not significant***




## Canopy volume vs height (ANCOVA)



```{r}

final.canopy.height.fit <- lmer(log.canopy.V ~ log.tree.H + rainfall + (1|landform), data = canopy_volume, REML = T)

summary(final.canopy.height.fit)

```


### assumptions
```{r}
qqnorm(residuals(final.canopy.height.fit ), pch = 1, frame = FALSE)
qqline(residuals(final.canopy.height.fit ), col = "steelblue", lwd = 2)

```


```{r}
scatter.smooth(residuals(final.canopy.height.fit ) ~ fitted(final.canopy.height.fit ))
```


```{r}
# Extract the residuals from the model
cv.residuals <- residuals(final.canopy.height.fit )

# Perform the ACF analysis
 acf(cv.residuals)
```



```{r}
vif(final.canopy.height.fit)
``
```







### dredging 


```{r}
# significance of rainfall

final.canopy.height.fit <- lmer(log.canopy.V ~ log.tree.H + rainfall + (1|landform), data = canopy_volume, REML = F)
final.canopy.height.null <- lmer(log.canopy.V ~ log.tree.H + (1|landform), data = canopy_volume, REML = T)

anova(final.canopy.height.null,final.canopy.height.fit,test = "LRT")

```

```{r}
# significance of height

final.canopy.height.fit <- lmer(log.canopy.V ~ log.tree.H + rainfall + (1|landform), data = canopy_volume, REML = F)
final.canopy.height.null <- lmer(log.canopy.V ~ rainfall + (1|landform), data = canopy_volume, REML = T)

anova(final.canopy.height.null,final.canopy.height.fit,test = "LRT")

```

### structural sensitivity analysis


```{r}
# significance of interaction effect

final.canopy.height.fit <- lmer(log.canopy.V ~ log.tree.H * rainfall + (1|landform), data = canopy_volume, REML = F)
final.canopy.height.null <- lmer(log.canopy.V ~ log.tree.H + rainfall + (1|landform), data = canopy_volume, REML = T)

anova(final.canopy.height.null,final.canopy.height.fit,test = "LRT")

```

***do not include interactive effect***



### hypothesis testing for ranodm effect

```{r}
#test if we need landform as a random effect

nBoot <- 1000
lrStat <- rep(NA, nBoot)
final.canopy.height.null <- aov(log.canopy.V ~ log.tree.H + rainfall, data = canopy_volume) #null w/o random effect
final.canopy.height.fit <- lmer(log.canopy.V ~ log.tree.H + rainfall + (1|landform), data = canopy_volume, REML = F) # full model
lrObs <- 2 * logLik(final.canopy.height.fit) - 2 * logLik(final.canopy.height.null) # observed test stat
for (iBoot in 1:nBoot)
{
  canopy_volume$logCanopySim <- unlist(simulate(final.canopy.height.null)) # resampled data
  bNull <- aov(logCanopySim ~ log.tree.H + rainfall, data = canopy_volume) # null model
  bAlt <- lmer(logCanopySim ~ log.tree.H + rainfall + (1 |landform), data = canopy_volume, REML = F) # full model
  lrStat[iBoot] <- 2 * logLik(bAlt) - 2 * logLik(bNull) # resampled test stat
}
mean(lrStat > lrObs) # P-value for test of landform effect
```
***no evidence to include landform***

## Final Canopy Volume vs Hight model


```{r}
final.canopy.height.aov <- aov(log.canopy.V ~ log.tree.H + rainfall, data = canopy_volume)

```



```{r}
vif(final.canopy.height.aov)
```



```{r}
# Extract the residuals from the model
cv.residuals <- residuals(final.canopy.height.aov)

# Perform the ACF analysis
 acf(cv.residuals)
```

```{r}
final.canopy.height.aov <- aov(log.canopy.V ~ log.tree.H * rainfall, data = canopy_volume)

Anova(final.canopy.height.aov, type="III") 

```

***non significant interaction = no change in slope***

```{r}
final.canopy.height.aov <- aov(log.canopy.V ~ log.tree.H + rainfall, data = canopy_volume)

Anova(final.canopy.height.aov, type="III") 

TukeyHSD(final.canopy.height.aov, "rainfall")
```

```{r}
canopyvsheight <- canopy_volume%>%
  ggplot( aes( tree_height, canopy_volume * 1000000, colour = rainfall))+
  geom_point(aes(shape = landform))+
  geom_smooth(method = "nls",
              formula = y ~b +x^m,
              se = F,
              method.args = list(start = list(m=3, b=0)))+
  ggtitle("b")+
  xlab("height (m)")+
  ylab(bquote("canopy volume  " (cm^3)))+
  theme_cowplot()
  theme(legend.position = "none",
        axis.title=element_text(size=12))


print(canopyvsheight)
```



## Canopy volume vs basal area (ANCOVA)



```{r}

final.canopy.basal.fit <- lmer(log.canopy.V ~ log.basal_area + rainfall + (1|landform), data = canopy_volume, REML = T)

summary(final.canopy.basal.fit)
```


### assumptions 

```{r}
qqnorm(residuals(final.canopy.basal.fit ), pch = 1, frame = FALSE)
qqline(residuals(final.canopy.basal.fit ), col = "steelblue", lwd = 2)

```


```{r}
scatter.smooth(residuals(final.canopy.basal.fit ) ~ fitted(final.canopy.basal.fit ))
```


```{r}
# Extract the residuals from the model
cb.residuals <- residuals(final.canopy.basal.fit )

# Perform the ACF analysis
 acf(cb.residuals)
```

```{r}
vif(final.canopy.basal.fit)
```



### dredging


```{r}

final.canopy.basal.fit <- lmer(log.canopy.V ~ log.basal_area + rainfall + (1|landform), data = canopy_volume, REML = F)

final.canopy.basal.null <- lmer(log.canopy.V ~ log.basal_area + (1|landform), data = canopy_volume, REML = F)

anova(final.canopy.basal.null,final.canopy.basal.fit,test = "LRT")
```

```{r}
final.canopy.basal.null <- lmer(log.canopy.V ~ rainfall + (1|landform), data = canopy_volume, REML = F)

anova(final.canopy.basal.null,final.canopy.basal.fit,test = "LRT")
```
***rainfall not significant!!***






```{r}

basalvscanopy <- canopy_volume%>%
  ggplot( aes( basal_area, canopy_volume, colour = rainfall))+
  geom_point(aes(shape = landform))+
     geom_smooth(method = "nls",
              formula = y ~ b + x^m,
              se = FALSE,
              method.args = list(start = list(b= 0.4660558, m = 1.13)))+
  ylab(bquote("canopy volume " (m^3)))+
  xlab(bquote("basal area " (cm^2)))+
  ggtitle("c")+
  theme_cowplot()+
  theme(axis.title=element_text(size=12,face="bold"))+
  theme(legend.position = "none")
  


 print(basalvscanopy)
```






# PLOT LEVEL METRICS 

## build data set

sub_plot_area, plots, plot_density, rainfall,landform

```{r}

#create subplot level statistics    
  plot_D2h <- D2H_model%>%
  group_by( subplot )%>%
    summarise(
      sub_plot_area = mean(sub_plot_area),
      subplot_density = mean(subplot_density),
      ave.biomass = mean(biomass),
      diameter = mean(diamiter),
      tot.biomass = sum(biomass),
      stem_area = mean(3.1415/4*diamiter^2),
      biomass.m2 = tot.biomass/sub_plot_area,
      plots = first(plots),
      plot_density = mean(plot_density),
      rainfall = first(rainfall),
      landform = first(landform))%>%
  ungroup()%>%
  mutate(
    logbiomass.m2 = log(biomass.m2),
    log.sub.dens = log(subplot_density))

summary(plot_D2h)

```


```{r}
stand_metrics <- plot_D2h %>% 
  group_by( subplot, landform, sub_plot_area, plots, rainfall, subplot_density, plot_density)%>%
  summarise(ave.biomass = mean(ave.biomass),
            logbiomass.m2 = mean(logbiomass.m2),
            log.sub.dens = mean(log.sub.dens),
            biomass.m2 = mean(biomass.m2))
  


summary(stand_metrics)
```


```{r}
#read in spatial co-ordinates 

subplot_coordinates <- read_csv("locational_data/subplot_coordinates.csv") #GPS centroid for each plot (WSG84).

summary(subplot_coordinates)

summary(stand_metrics)

```


```{r}

# bind the each co-ordinate to the relevant plot 

coords <- subplot_coordinates%>%
  rename(subplot = sub_name )%>%
  mutate(subplot = as.factor(subplot))


spatial_stand_metrics <- cbind(stand_metrics,coords)

# class xcoord and ycoord as coordinates using the coordinates() function 

coordinates(spatial_stand_metrics) <- ~ xcoord + ycoord

summary(spatial_stand_metrics)
```







## Plot level biomass (LMEM) 



```{r}

#compare main effects to interactive model 

final.plot.biomass.fit <- lmer(logbiomass.m2 ~ rainfall + (1|landform/plots), data = stand_metrics, REML = T)

summary(final.plot.biomass.fit)

```


### assumptions

```{r}
qqnorm(residuals(final.plot.biomass.fit), pch = 1, frame = FALSE)
qqline(residuals(final.plot.biomass.fit), col = "steelblue", lwd = 2)

```


```{r}
scatter.smooth(residuals(final.plot.biomass.fit) ~ fitted(final.plot.biomass.fit))
```


```{r}
# Extract the residuals from the model
residuals <- residuals(final.plot.biomass.fit)

# Perform the ACF analysis
 acf(residuals)
```

### dredging 

```{r}
final.plot.biomass.fit <- lmer(logbiomass.m2 ~ rainfall +  (1|landform/plots), data = stand_metrics, REML = F)

final.plot.biomass.null <- lmer(logbiomass.m2 ~ (1|landform/plots), data = stand_metrics, REML = F)

anova(final.plot.biomass.null,final.plot.biomass.fit,test = "LRT")
```


***evidence to include rainfall***

### hypothesis testing for random effects



final.plot.biomass.fit lmer(logbiomass.m2 ~ rainfall +  (1|landform/plots), data = stand_metrics, REML = F)

```{r}
#test if we need landform given we have plots as a random effect

nBoot <- 1000
lrStat <- rep(NA, nBoot)
final.plot.biomass.null <- lmer(logbiomass.m2 ~ rainfall +  (1|plots), data = stand_metrics, REML = F) #null w/o random effect
final.plot.biomass.fit <- lmer(logbiomass.m2 ~  rainfall + (1|landform/plots), data = stand_metrics, REML = F) # full model
lrObs <- 2 * logLik(final.plot.biomass.fit) - 2 * logLik(final.plot.biomass.null) # observed test stat
for (iBoot in 1:nBoot)
{
  stand_metrics$logBiomassSim <- unlist(simulate(final.plot.biomass.null)) # resampled data
  bNull <- lmer(logBiomassSim ~ rainfall +  (1|plots), data = stand_metrics, REML = F) # null model
  bAlt <- lmer(logBiomassSim ~  rainfall + (1|landform/plots), data = stand_metrics, REML = F) # full model
  lrStat[iBoot] <- 2 * logLik(bAlt) - 2 * logLik(bNull) # resampled test stat
}
mean(lrStat > lrObs) # P-value for test of landform effect
```



***no evidence to include landform***


```{r}
#test if we need plots as a random effect

nBoot <- 1000
lrStat <- rep(NA, nBoot)
final.plot.biomass.null <- aov(logbiomass.m2 ~ rainfall, data = stand_metrics) #null w/o random effect
final.plot.biomass.fit <- lmer(logbiomass.m2 ~  rainfall + (1|plots), data = stand_metrics, REML = F) # full model
lrObs <- 2 * logLik(final.plot.biomass.fit) - 2 * logLik(final.plot.biomass.null) # observed test stat
for (iBoot in 1:nBoot)
{
  stand_metrics$logBiomassSim <- unlist(simulate(final.plot.biomass.null)) # resampled data
  bNull <- aov(logBiomassSim ~ rainfall, data = stand_metrics) # null model
  bAlt <- lmer(logBiomassSim ~  rainfall + (1|plots), data = stand_metrics, REML = F) # full model
  lrStat[iBoot] <- 2 * logLik(bAlt) - 2 * logLik(bNull) # resampled test stat
}
mean(lrStat > lrObs) # P-value for test of plots effect
```

***evidence to plots and a random effect*** 


## Final Plot level biomass Model

```{r}
final.plot.biomass.fit <- lmer(logbiomass.m2 ~  rainfall + (1|plots), data = stand_metrics, REML = F)

summary(final.plot.biomass.fit)
```

```{r}
pairs(emmeans(final.plot.biomass.fit, ~  rainfall))
```

***sig diff between moderate and high, approaching significance (0.0566) for low and high***


***model summary suggests the high rainfall category is different to the low rainfall category, but the estimated marginal means indicates that it is not?***

```{r}
stands_boimass <-stand_metrics %>%
  ggplot(aes( rainfall, biomass.m2, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab(bquote("biomass "(kg ~ m^-2)))+
  xlab( " d2h ")+
  ggtitle("a")+
  theme_cowplot()+
  theme(legend.position = "none")

print(stands_boimass)
```




## Stand density model (LMEM)


```{r}
final.density.fit <- lmer(log.sub.dens ~ rainfall + (1|landform/plots), data = stand_metrics, REML = F)

summary(final.density.fit)
```


### assumptions


```{r}
qqnorm(residuals(final.density.fit), pch = 1, frame = FALSE)
qqline(residuals(final.density.fit), col = "steelblue", lwd = 2)

```


```{r}
scatter.smooth(residuals(final.density.fit) ~ fitted(final.density.fit))
```


```{r}
# Extract the residuals from the model
dens.residuals <- residuals(final.density.fit)

# Perform the ACF analysis
 acf(dens.residuals)
 
```


### dreddging 


```{r}
final.density.fit <- lmer(log.sub.dens ~ rainfall +  (1|landform/plots), data = stand_metrics, REML = F)

final.density.null <- lmer(log.sub.dens ~ (1|landform/plots), data = stand_metrics, REML = F)

anova(final.density.fit,final.density.null,test = "LRT")
```

***rainfall is significant***

### hypothesis testing



```{r}
#test if we need landform given we have plots as a random effect

nBoot <- 1000
lrStat <- rep(NA, nBoot)
final.density.null <- lmer(log.sub.dens ~ rainfall +  (1|plots), data = stand_metrics, REML = F) #null w/o random effect
final.density.fit <- lmer(log.sub.dens ~  rainfall + (1|landform/plots), data = stand_metrics, REML = F) # full model
lrObs <- 2 * logLik(final.density.fit) - 2 * logLik(final.density.null) # observed test stat
for (iBoot in 1:nBoot)
{
  stand_metrics$logdensitySim <- unlist(simulate(final.density.null)) # resampled data
  bNull <- lmer(logdensitySim ~ rainfall +  (1|plots), data = stand_metrics, REML = F) # null model
  bAlt <- lmer(logdensitySim ~  rainfall + (1|landform/plots), data = stand_metrics, REML = F) # full model
  lrStat[iBoot] <- 2 * logLik(bAlt) - 2 * logLik(bNull) # resampled test stat
}
mean(lrStat > lrObs) # P-value for test of landform effect
```


***no evidence to include landforms as a randome effect***



```{r}
#test if we need plots as a random effect

nBoot <- 1000
lrStat <- rep(NA, nBoot)
final.density.null <- aov(log.sub.dens ~ rainfall, data = stand_metrics) #null w/o random effect
final.density.fit <- lmer(log.sub.dens ~ rainfall +  (1|plots), data = stand_metrics, REML = F) # full model
lrObs <- 2 * logLik(final.density.fit) - 2 * logLik(final.density.null) # observed test stat
for (iBoot in 1:nBoot)
{
  stand_metrics$logdensitySim <- unlist(simulate(final.density.null)) # resampled data
  bNull <- aov(log.sub.dens ~ rainfall, data = stand_metrics) # null model
  bAlt <- lmer(logdensitySim ~  rainfall + (1|plots), data = stand_metrics, REML = F) # full model
  lrStat[iBoot] <- 2 * logLik(bAlt) - 2 * logLik(bNull) # resampled test stat
}
mean(lrStat > lrObs) # P-value for test of plots effect
```




### change model structure to ANCOVA 

```{r}
final.density.aov <- aov(log.sub.dens ~ rainfall, data = spatial_stand_metrics)
```


### assumptions



```{r}
qqnorm(residuals(final.density.aov), pch = 1, frame = FALSE)
qqline(residuals(final.density.aov), col = "steelblue", lwd = 2)

```


```{r}
scatter.smooth(residuals(final.density.aov) ~ fitted(final.density.aov))
```

```{r}
# Extract the residuals from the model
dens.residuals.aov <- residuals(final.density.aov)

# Perform the ACF analysis
 acf(dens.residuals.aov)
 
```


***test if autocorelation is significant***

```{r}
spatial_stand_metrics$lagged_density <- lag(stand_metrics$log.sub.dens, 1)
```

```{r}
final.density.lag.fit <- aov(log.sub.dens ~ rainfall + lagged_density, data = spatial_stand_metrics)

Anova(final.density.lag.fit)

coef(final.density.lag.fit)
```

***evidence of significant autocorrelation*** 


### check for spatial_autocorrelation


```{r}

#add the residuals to the data set

spatial_stand_metrics$Resid <- residuals(final.density.aov)


bubble(spatial_stand_metrics, "Resid")

```


```{r}
coords = cbind(spatial_stand_metrics$xcoord, spatial_stand_metrics$ycoords)
w = fields:::rdist(coords)
Moran.I(x = spatial_stand_metrics$log.sub.dens, w = w)
```

***not spatial auto_correlations***

***despite the non-significance of plots as a random effect as determined by bootstrapping, the inclusion of this variable eliminates significant autocorrelation and therefore is included to allow the data to meet the assumption of dependence in linear modeling***


## Final stand density model

```{r}
final.density.fit <- lmer(log.sub.dens ~ rainfall + (1|plots), data = stand_metrics, REML = F)


pairs(emmeans(final.density.fit, ~ rainfall))
```




```{r}
# Extract the residuals from the model
dens.residuals.fit <- residuals(final.density.fit)

# Perform the ACF analysis
 acf(dens.residuals.fit)
 
```


## Average tree biomass VS tree density (LMEM)

```{r}

final.add.competition.model <- lmer(log.sub.dens ~ log(ave.biomass) + rainfall + (1|landform /plots), data = stand_metrics, REML = T)

summary(final.add.competition.model)
```

### assumptions

```{r}
vif(final.add.competition.model)
```


```{r}
biomass_rainfall <- plot_D2h %>%
  ggplot(aes( rainfall, ave.biomass, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab("subplot mean biomass")+
  xlab( " d2h ")+
  ggtitle("a")+
  theme_cowplot()+
  theme(legend.position = "none")

print(biomass_rainfall)
```













