---
title: "SFA"
output: html_notebook
---

## SFA



```{r}

#create frontier model. 
self.thin <- sfa(log10(ave.biomass) ~ log10(subplot_density) * rainfall ,  data = plot_D2h)

summary(self.thin, extraPar = T)

# testing performance of model against OLS

lrtest(self.thin)

```
# sigmaU coeeficeint indicates signifcance of the inneficency term. error decomposition means means the decomosition of the random error term into error and ineficency 


#### assumptions

```{r}
scatter.smooth(residuals(self.thin) ~ fitted(self.thin))
```

```{r}
qqnorm(residuals(self.thin), pch = 1, frame = FALSE)
qqline(residuals(self.thin), col = "steelblue", lwd = 2)
```

```{r}
# Extract the residuals from the model
residuals <- residuals(self.thin)

# Perform the ACF analysis
 acf(residuals)


```




# combined approach 


self.thin <- sfa(log10(biomass) ~ log10(subplot_density) * rainfall  ,  data = plot_D2h)


## step 1 FIT LMEM WITH ONLY RANDOM EFFECTS
```{r}

self.thin.lme <- lmer(biomass ~  (1|subplot_density) * (1|rainfall) * (1|landform) + (1|plots), data = plot_D2h,  REML = T)

summary(self.thin.lme)


```

```{r}
scatter.smooth(residuals(self.thin.lme) ~ fitted(self.thin.lme))
```






## STEP 2 EXTRACT BLUP and ADD to data set




```{r}

#CALCULATE ESTIMATES AND VARIANCE FOR RANDOM EFFECTS
STvarComp <- as.data.frame(VarCorr(self.thin.lme,comp = "vcov")) 

#extract coeffient estimates for random effects

ST.ran <- ranef(self.thin.lme)$plots
ST.ran <- setNames(ST.ran, c("BLUPs"))

#label plot numbers
ST.ran<- ST.ran %>%
  mutate(plots = 1:19)


ST.stem_plots <- merge(plot_D2h, ST.ran, by = "plots", all.x = TRUE)

print(ST.stem_plots)

```





## step 3 incorpurate random effects into SFA

```{r}
sfa_self_thin <- sfa(log10(biomass) ~ log10(subplot_density) * rainfall + landform + BLUPs, data = ST.stem_plots)

summary(sfa_self_thin, extraPar = T)

#compare to ordinary least squares 


lrtest(sfa_self_thin)
```

## step 4 check assumptions
```{r}
scatter.smooth(residuals(sfa_self_thin) ~ fitted(sfa_self_thin))
qqnorm(residuals(sfa_self_thin), pch = 1, frame = FALSE)
qqline(residuals(sfa_self_thin), col = "steelblue", lwd = 2)
```
## step 5 compare effeicncy scores against normal SFA approach. 





## maximizing biomass production

```{r}

#create frontier model. rainfall and landform are exogenous variables with landform nested within rainfall

max.m2 <- sfa(log10(biomass.m2) ~ log10(subplot_density) * rainfall + landform, data = plot_D2h)

summary(max.m2, extraPar = T)

# testing performance of model against OLS

lrtest(max.m2)

```


```{r}

scatter.smooth(residuals(max.m2) ~ fitted(max.m2))

```

```{r}
# Extract the residuals from the model
residuals <- residuals(max.m2)

# Perform the ACF analysis
 acf(residuals)


```



```{r}
qqnorm(residuals(max.m2), pch = 1, frame = FALSE)
qqline(residuals(max.m2), col = "steelblue", lwd = 2)
```









## ploting frontiers

```{r}
# plotted self thining

ggplot(stem_plots, aes(log(subplot_density), log(ave.AGB), colour = rainfall)) +
  geom_point(aes(shape = factor(landform)))+
  geom_abline(intercept = 3.147408, slope = 0.020039, linetype = "solid", color = "red", size = 1) +
  geom_abline(intercept = 3.089888, slope = -0.456046, linetype = "solid", color = "dark green", size = 1) +
  geom_abline(intercept = 0.70  , slope =   -1.1242880, linetype = "solid", color = "blue", size = 1)+
  theme_cowplot()+
  labs(x = "Subplot Density", y = "average tree biomass", color = "Rainfall", shape = "Landform")
  


```

```{r}
#change self thinning line plot scale to log 

#start by defining the functions for each slope. us the coefficients in the above outputs.For raise e to the power of the intercept term to find normal scale intercept.Raise x to the slope to term define the slope of the curve. Functions are expected to be defined as a function of x and not as a formula.

low_thin.fun <- function(x) {3.781* x^ -0.26900772}
mod_thin.fun <- function(x) {21 * x^-0.456046}
high_thin.fun <- function(x) {2.23* x^ -0.8166378}




ggplot(stem_plots, aes(subplot_density,ave.AGB, colour = rainfall)) +
  geom_point(aes(shape = factor(landform)))+
  scale_y_log10()+
  scale_x_log10()+
  geom_function(fun = low_thin.fun, n = 127, colour = "red", size = 1)+
  geom_function(fun = mod_thin.fun, n = 127, colour = "dark green", size = 1)+
  geom_function( fun = high_thin.fun, n = 127, colour = "blue", size = 1)+
    #ggtitle("c")+
  ylab( bquote("mean tree biomass " (kg)))+
  xlab( bquote( "tree density " (tree ~ m^-2)))

  

  
```

```{r}
ggplot(plot_D2h, aes(biomass,subplot_density, colour = rainfall)) +
  geom_point(aes(shape = factor(landform)))+
  scale_y_log10()+
  scale_x_log10()+
  geom_function(fun = low_thin.fun, n = 127, colour = "red", size = 1)+
  xlab( bquote("mean biomass " (kg)))+
  ylab( bquote( "tree density " (tree ~ m^-2)))
```







```{r}
#normal scale using function determined in the above chunk. 

self.thining <- ggplot(stem_plots, aes(subplot_density, ave.AGB, colour = rainfall)) +
  geom_point(aes(shape = factor(landform)))+
  theme_cowplot()+
  scale_x_log10()+
  scale_y_log10()+
  ylab( bquote("mean tree biomass " (kg)))+
  xlab( bquote( "tree density " (tree ~ m^-2)))+
  geom_function(fun = low_thin.fun, n = 127, colour = "red", size = 1)+
  geom_function(fun = mod_thin.fun, n = 127, colour = "dark green", size = 1)+
  geom_function( fun = high_thin.fun, n = 127, colour = "blue", size = 1)+
  ggtitle("c")+
  theme(axis.title=element_text(size=12),legend.position = "none")

print(self.thining)
```




```{r}
# plotted maximum biomass efficiency frontiers

ggplot(stem_plots, aes(log(subplot_density), log(biomass.m2), colour = rainfall)) +
  geom_point(aes(shape = factor(landform))) +
  geom_abline(intercept = 2.268333, slope = 0.326692, linetype = "solid", color = "red", size = 1) +
  geom_abline(intercept = 3.894589, slope = 0.808454, linetype = "solid", color = "dark green", size = 1) +
  geom_abline(intercept = 1.750087  , slope = -0.088450 , linetype = "solid", color = "blue", size = 1) +
  labs(x = "Subplot Density", y = "Biomass/m2", color = "Rainfall", shape = "Landform") +
  theme_cowplot()
 

```

```{r}
#change maximum potential biomass plot scale to log 

#start by defining the functions for each slope. us the coefficients in the above outputs.For raise e to the power of the intercept term to find normal scale intercept.Raise x to the slope to term define the slope of the curve. Functions are expected to be defined as a function of x and not as a formula.

low_max.fun <- function(x) {9.71 * x^0.3628}
mod_max.fun <- function(x) {53.76 * x^0.808454}
high_max.fun <- function(x) {5.0869 * x^ -0.1243}




max.potent <- ggplot(stem_plots, aes(subplot_density,biomass.m2, colour = rainfall)) +
  geom_point(aes(shape = factor(landform)))+
  scale_y_log10()+
  scale_x_log10()+
  geom_function(fun = low_max.fun, n = 127, colour = "red", size = 1)+
  geom_function(fun = mod_max.fun, n = 127, colour = "dark green", size = 1)+
  geom_function( fun = high_max.fun, n = 127, colour = "blue", size = 1)+
    #ggtitle("c")+
  ylab( bquote("plot biomass " (kg ~ m^-2)))+
  xlab( bquote( "tree density " (tree ~ m^-2)))+
  theme_cowplot()+
  ggtitle("d")+
  theme(axis.title=element_text(size=12),
        legend.position = "none")
print(max.potent)

```


```{r}

ggplot(stem_plots, aes(subplot_density, biomass.m2, colour = rainfall)) +
  geom_point(aes(shape = factor(landform)))+
  theme_cowplot()+
  geom_function(fun = low_max.fun, n = 127, colour = "red", size = 1)+
  geom_function(fun = mod_max.fun, n = 127, colour = "dark green", size = 1)+
  geom_function( fun = high_max.fun, n = 127, colour = "blue", size = 1)+
    #ggtitle("c")+
  ylab( bquote("plot biomass " (kg ~ m^-2)))+
  xlab( bquote( "tree density " (tree ~ m^-2)))+
  xlim(0, 0.25)+
  ylim(0, 10)
```