---
title: "R Notebook"
output: html_notebook
---
# set up



```{r}


library(forecast)

library('tseries')

library(rio)

library(lme4)

library(emmeans)

library(tidyverse)

library(readr)

library(nlme)

library(REAT)

library(ggpmisc)

library(cowplot)
```


# CHIRPS rainfall Data
## load data set 
```{r}
rainfall_history <- read_csv("rainfall_data/rainfall_history.csv")


an.rainfall <- rainfall_history%>%
  mutate(
  rainfall = as.factor(rainfall),
  rainfall = factor(rainfall, levels = c("low", "moderate", "high")))

f.an.rainfall <- an.rainfall%>%
  filter(year > 1985)

summary(an.rainfall)
```


## plot chirps data

```{r}
rainfallTS<- f.an.rainfall%>%
  ggplot( aes(year, annual_precipitation, colour = rainfall))+
  geom_line()+
    ylab("annual precipitation (mm)")+
    xlab("year")+
    ggtitle(" ")+
  theme_minimal()
  
  
print(rainfallTS) 
#ggsave("rainfallTS.png")

```


```{r}

an.rainfall%>%
  ggplot(aes(rainfall, annual_precipitation, colour = rainfall))+
  geom_boxplot()
```
```{r}

  
rainfall.avo  <- aov(annual_precipitation ~ rainfall, data = an.rainfall)

summary(rainfall.avo)

TukeyHSD(rainfall.avo)
```

```{r}
qqnorm(residuals(rainfall.avo), pch = 1, frame = FALSE)
qqline(residuals(rainfall.avo), col = "steelblue", lwd = 2)

```


```{r}
plot(rainfall.avo, which = 1)
```


## unpack variation for each site and compare

### homogenity of variance 
```{r}

bartlett.test(annual_precipitation ~ rainfall, data = an.rainfall)

```


### coeficent of variation is a measure of variance of a dataset that takes scale into account 

```{r}
high.cv <- apply(an.rainfall[1:40,3], MARGIN = 2, FUN = cv)

moderate.cv<- apply(an.rainfall[41:80,3], MARGIN = 2, FUN = cv)

low.cv <- apply(an.rainfall[81:120,3], MARGIN = 2, FUN = cv)

```





# bom rainfall data 

```{r}

bom_rainfall <- read_csv("rainfall_data/bom/total_bom.csv")



summary(bom_rainfall)

```


## plot data for each site. facet wrap

```{r}
bom_rainfall%>%
  ggplot( aes(Year, Annual, colour = station_ID))+
  geom_line()+
  facet_wrap( ~ site)
  
  

```



# compare ndvi across sites and landforms  
```{r}

total_bands <- read_csv("NDVI_DATA/total_bands.csv")


```


## calculate annual avearage NDVI for each site and landform

```{r}

 total_nir <- total_bands%>%
  aggregate(NIR ~ year + site + landform, mean)

total_red <- total_bands%>%
  aggregate(Red ~ year + site + landform, mean)

total_red_nir <- left_join(total_red, total_nir, by = c("year", "site", "landform"))

total_ndvi <- total_red_nir%>%
  mutate(
    NDVI =  (NIR-Red)/(NIR+Red),
    site = factor(site, levels = c("low", "moderate", "high")))

```



## visualise with site and landform for each NDVI time series 

```{r}

v <- rep(0, 100)
v{c()}


NDVI_TS<- total_ndvi%>%
  ggplot( aes(year,NDVI, colour = site, shape = landform))+
  geom_line()+
  geom_point(aes(shape=factor(landform))) + 
  labs(shape = "landform", colour = "site")+
  ylab("NDVI value")+
  xlab("year")+
  ggtitle("b")+
  theme_minimal()
 
  
print(NDVI_TS)

#ggsave("NDVI_TS.png")

```

***note how nocoleche's accumulation sites NDVI values return to to normal after the millennium drought, whereas all other sites do***
***also no how nocoleche received simialr rainfall to 


```{r}

```


# dieback tree size comparison

```{r}
library(readr)
noco_deadvalive <- read_csv("rainfall_data/noco_deadvalive.csv")

```
## construct data set 

```{r}
noco_basal <- noco_deadvalive%>%
  gather("d_10", "diamiter", starts_with ("d_"))%>%
  mutate(
    basal_area = (diamiter*0.5)^2*3.14159265359,
    log_basal = log(basal_area))%>%
  arrange(tree)%>%
  filter(!is.na(diamiter))
```

## compare living and dead basal area visually 
```{r}
noco_basal%>%
  ggplot(aes(basal_area,colour = stem))+
  geom_histogram()+
  scale_x_log10()+
  theme_grey()

```


```{r}

dead_vs_alive <- noco_basal%>%
  ggplot(aes(status, basal_area, fill = status))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab(bquote("basal area "(cm^2)))+
  xlab("tree status")+
  #ggtitle("dead vs alive")+
  theme_cowplot()

 
print(dead_vs_alive)    
ggsave("dead_vs_alive.png")
```



## t-test 

```{r}
t.test(log_basal ~ status, data = noco_basal, var.equal = TRUE)

```
```{r}
dead.tree <- filter(noco_basal, status == "d")

alive.tree <- filter(noco_basal, status == "a")

mean(dead.tree$basal_area)
mean(alive.tree$basal_area)

```

## check anova assumptions

```{r}
qqnorm(noco_basal$log_basal, pch = 1, frame = FALSE)
qqline(noco_basal$log_basal, col = "steelblue", lwd = 2)
```





