---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# plots


# set up

```{r}

library(lme4) #mixed effect linear models

library(lmerTest)

library(emmeans)

library(tidyverse)

library(readr)

library(nlme)

library(ggpmisc)

library(cowplot)

library(car)

library(forcats)

library(mgcv)

library(glmmLasso)

library (glmmTMB)

```

```{r}
total_plots <- read_csv("plot_data/total_plots.csv")%>%
  mutate(
   landform = as.factor(landform),

    plots = as.factor(plots),
   
   subplot = as.factor(subplot),

    rainfall = factor(rainfall, levels = c("low", "moderate" , "high")))

summary(total_plots)
```
# build data set

# calcuale basal area and summed biomass



```{r}

 #calculate basal area   
  basal_area <- total_plots %>%
  gather("d_10", "diamiter", starts_with ("d_"))%>%
  mutate(basal_area = (diamiter*0.5)^2*3.14159265359)%>%
  arrange(tree)%>%
  filter(!is.na(diamiter))
```

#d2h model


M(kg) = 34.4703 - 8.0671 DBH + 0.6589 DBH^2 (cm)

```{r}
  
 # using the accacia allometric equation from feller 1980 we will calculate biomass for each tree based on a "stem by stem" analysis.
# see "Review of Allometric Relationships for Estimating Woody Biomass for New South Wales, the Australian Capital Territory, Victoria, Tasmania and South Australia" (page 36)
    D2H_model <- basal_area %>%
   group_by(tree, height, landform, plots, rainfall, subplot, sub_plot_area, subplot_density, plot_density)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
      mutate(diamiter = (sqrt(basal_area/3.14159265359))*2,
             dbh = (diamiter*0.798)-0.577,
             stem_log_biomass = log(1.9+ 424.9 *(dbh/100)^2 * (height/100)),
             stem_biomass = exp(stem_log_biomass),
             branch_log_biomass = log(208) + log((dbh/100)^2 * (height/100)),
             branch_biomass = exp(branch_log_biomass),
             leave_log_biomass = log(8) + log((dbh/100)^2 * (height/100)),
             leave_biomass = exp(leave_log_biomass))%>%
      mutate(
        biomass = stem_biomass + branch_biomass + leave_biomass,
        log_biomass = log(biomass))

```
    

#d2h plot level biomass

```{r}

    
  plot_D2h <- D2H_model%>%
  group_by( subplot, landform, sub_plot_area, plots, rainfall, subplot_density, plot_density)%>%
    summarise(
      ave.biomass = mean(biomass),
      diameter = mean(diamiter),
      tot.biomass = sum(biomass),
      stem_area = mean(3.1415/4*diamiter^2),
    biomass.m2 = tot.biomass/sub_plot_area)%>%
  ungroup()%>%
  mutate(
    logbiomass.m2 = log(biomass.m2),
    log.sub.dens = log(subplot_density),
    plots = as.numeric(plots))%>%
  ungroup()

summary(plot_D2h)

```






```{r}
names(plot_D2h)
```


```{r}
stand_metrics <- plot_D2h %>% 
  group_by( subplot, landform, sub_plot_area, plots, rainfall, subplot_density, plot_density)%>%
  summarise(ave.biomass = mean(ave.biomass),
            logbiomass.m2 = mean(logbiomass.m2),
            log.sub.dens = mean(log.sub.dens),
            biomass.m2 = mean(biomass.m2))
  


summary(stand_metrics)
```









# diameter only model varied for tree structure. 

```{r}
#combind basal area for multi stemmed trees for a "combined stem analysis" 
 ind_biomass <- basal_area%>%
  group_by(tree, height, landform, plots, rainfall, subplot, sub_plot_area, subplot_density, plot_density)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
  mutate(
     #back tranform total_basal_area to diamiter 
    diamiter = case_when( rainfall == "low" ~ ((sqrt(basal_area/3.14159265359))*2),
                          rainfall == "moderate" ~ ((sqrt(basal_area/3.14159265359))*2),
                          rainfall == "high" ~ ((sqrt(basal_area/3.14159265359))*2)*0.798)-0.577, # converting d10 to d130
     # Equations taken from -> Paul et al, "testing the generality of above ground ind_biomass allometry across plant functional types at the continent scale" -> Fmulti "ln(AGB) = 2.474ln(D10)-2.757" used for lower rainfall sites (page 9)
    logAGB = case_when(rainfall == 'low' ~ 2.474*log(diamiter)-2.757,
                      rainfall == 'moderate' ~ 2.474*log(diamiter)-2.757,
                      rainfall == 'high' ~ 2.220*log(diamiter)-1.693),
    # correctional factors taken form relevant equations in respective papers. CF for high rainfall taken form average of Acacia trees paul and roxburgh 2013.
    AGB = case_when( rainfall == 'low' ~ exp(logAGB)*1.0787,
                     rainfall == 'moderate' ~ exp(logAGB)*1.0787,
                     rainfall == 'high' ~ exp(logAGB)*1.0436))
 
    
 #plot level biomass
plot_biomass <- ind_biomass%>%
  aggregate(AGB ~ subplot + landform + sub_plot_area + plots + rainfall + subplot_density + plot_density, sum)%>%
  mutate(
    biomass.m2 = AGB/sub_plot_area,
    logbiomass.m2 = log(biomass.m2),
    log.sub.dens = log(subplot_density))

```



#paul tree model 

```{r}
#combind basal area for multi stemmed trees for a "combined stem analysis" 
Paul_tree <- basal_area%>%
  group_by(tree, height, landform, plots, rainfall, subplot, sub_plot_area, subplot_density, plot_density)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
  mutate(
     #back tranform total_basal_area to diamiter 
    diamiter =  (((sqrt(basal_area/3.14159265359))*2)*0.798)-0.577,
    logAGB = 2.220*log(diamiter)-1.693,
    AGB = exp(logAGB)*1.0436)
    
 #plot level biomass
p.tree_plots <- Paul_tree%>%
  aggregate(AGB ~ subplot + landform + sub_plot_area + plots + rainfall + subplot_density + plot_density, sum)%>%
  mutate(
    biomass.m2 = AGB/sub_plot_area,
    logbiomass.m2 = log(biomass.m2),
    log.sub.dens = log(subplot_density))

```


# paul shrub model

```{r}

#individual level

paul_shrub <- basal_area%>%
  group_by(tree, height, landform, plots, rainfall, subplot, sub_plot_area, subplot_density, plot_density)%>%
  summarise(basal_area = sum(basal_area),.groups = "drop")%>%
  mutate( 
    diamiter = ((sqrt(basal_area/3.14159265359))*2),
    logAGB = 2.474*log(diamiter)-2.757,
    AGB = exp(logAGB)*1.0787)


#plot level 
    
p.shrub_plots <- paul_shrub%>%
  aggregate(AGB ~ subplot + landform + sub_plot_area + plots + rainfall + subplot_density + plot_density, sum)%>%
  mutate(
    biomass.m2 = AGB/sub_plot_area,
    logbiomass.m2 = log(biomass.m2),
    log.sub.dens = log(subplot_density))


```



```{r}
summary(Paul_tree)
```

```{r}
summary(paul_shrub)
```

```{r}
 summary(D2H_model)
```

```{r}
 summary(ind_biomass)
```




```{r}
isocline <- ind_biomass %>%
  group_by(subplot,  landform, sub_plot_area, plots , rainfall)%>%
  summarise(
    
    basal_area = sum(basal_area),
    subplot.n = n(),
    subplot_density = subplot.n / sub_plot_area[1],
    ave.AGB = mean(AGB),
    ) %>% ungroup()


isocline$rainfall <-  factor(isocline$rainfall, 
                             levels = c("low", "moderate" , "high"))


summary(isocline)

  
```


```{r}

ind_biomass%>%
  mutate(basal_area = as.numeric(basal_area))

summary(ind_biomass)

```



```{r}

summary(plot_biomass)

```



# filtering
```{r}
f.ind_biomass <- filter(ind_biomass, diamiter > 2)

f.D2H_model <- filter(D2H_model, diamiter > 2)

f.paul_shrub <- filter(paul_shrub, diamiter >2)

f.paul_tree <- filter(Paul_tree, diamiter >2)

```



```{r}
d2h.h.rain. <- filter(f.D2H_model, rainfall == "high")
diam.h.rainf <- filter(ind_biomass, rainfall == "high")
```




```{r}

#plot damiter measurements 

#diamiter
diam.h.rainf%>%
ggplot( aes(x= AGB)) +
  geom_histogram()+
  ggtitle("diameter only")
  

d2h.h.rain.%>%
  ggplot( aes(x = biomass))+
  geom_histogram()+
  ggtitle("d2h model")
  
```





```{r}

#plot biomass measurements 

#biomass
f.ind_biomass%>%
ggplot(aes(x=AGB, colour = factor(rainfall))) +
  geom_histogram( fill = "#69b3a2",
                  alpha = 0.09,
                  bins = 40)+
  ggtitle("summed basal")+
  scale_x_log10()+
  theme_cowplot()+
  theme(plot.title = element_text(size = 15))

#D2h biomass model

f.D2H_model %>%
  ggplot( aes( x = biomass, colour = factor (rainfall)))+
   geom_histogram( fill = "#69b3a2",
                  alpha = 0.09,
                  bins = 40)+
  scale_x_log10()+
  ggtitle("stem by stem ")+
  theme_cowplot()+
  theme(plot.title = element_text(size = 15))
```


```{r}

#plot the ind_biomass dist

#multistem

ind_biomass%>%
ggplot( aes(x=AGB, colour = factor(rainfall))) +
  geom_histogram( fill = "#69b3a2",
                  alpha = 0.09,
                  bins = 40)+
  scale_x_log10()+
  ggtitle("log scale ind_biomass")+
  theme_classic()+
  theme(plot.title = element_text(size = 15))

ind_biomass%>%
  ggplot( aes(x=AGB, colour = factor(rainfall)))+
            geom_histogram(bins = 40)+
            theme_grey()+
            ggtitle("Normal Scale")+
     
         theme(plot.title = element_text(size = 15))
```


# basal area and height 



```{r}

fb.d2H <- D2H_model%>%
  filter(basal_area < 1450)
summary(fb.d2H)

```

```{r}
basal.vs.heigh_plots <- fb.d2H%>%
  ggplot(aes(basal_area, height, colour = rainfall))+
  geom_point(aes(shape = landform))+
geom_smooth(method = "nls",
              formula = y ~ c*x^z,
             se = FALSE,
             method.args = list(start = list(c=125, z=0.383739)))+
  ylab("height (cm)")+
  xlab(bquote("basal area " (cm^2)))+
  ggtitle("a")+
  theme_cowplot()+
  theme(axis.title=element_text(size=12), legend.position = "none")
  

print(basal.vs.heigh_plots)

#ggsave("basal.vs.height_plots.png")

```





## likely power-law relationships 

## plot log(n) of height and ind_biomass and fit lm for high rainfall

```{R}

h.rain. <- filter(f.D2H_model, rainfall == "high")

h.model <- lm(log(height) ~ log(basal_area), data = h.rain.)

summary(h.model)
exp(4.829160)

```





***using intercep and slop form above (note good r-squared and p-value) calculate power law line of best fit (rais e to the power of said constants)***
***log height = 5.12975 + 0.307957*basal_area***
***height = 168.9752 + basal_area^0.307957*** 


```{r}
nlm.high <- nls(height ~ c*basal_area^z,
    data = h.rain.,
    start = c(c=168.9752, z =0.307957))

summary(nlm.high)

2.71828^5.12975
```

```{r}

h.rain.%>%
  ggplot(aes(basal_area, height))+
  geom_point()+
  theme_bw()+
  geom_smooth(method = "nls",
              formula = y ~ c*x^z,
             se = FALSE,
             method.args = list(start = list(c=168.9752, z=0.307957)))

```

## plot log(n) of height and ind_biomass and fit lm for low rainfall



```{r}
l.rain. <- filter(f.D2H_model, rainfall != "high")

l.model <- lm(log(height) ~ log(basal_area), data = l.rain.)

noco.se <- filter(l.rain., rainfall == "moderate")

sturt.se <- filter(l.rain., rainfall == "low")


summary(l.model)

exp(4.71186)
```




***use interecept and slope estimates to calcualte curve of best fit***
***log(height) = 4.71186 + 0.6386*log(basal_area)***
***height = 111.2589 + basal_area^0.28933***



```{r}
nlm.low <- nls(height ~ c*basal_area^z,
    data = l.rain.,
    start = c(c=111.2589, z = 0.28933))

summary(nlm.low)
```


```{r}

N.model <- lm(log(height) ~ log(basal_area), data = noco.se)

summary(N.model)

exp(4.373832)
```


```{r}

S.model <- lm(log(height) ~ log(basal_area), data = sturt.se)

summary(S.model)

exp(5.131581)
```






```{r}

l.rain.%>%
  ggplot(aes(basal_area, height))+
  geom_point()+
  theme_bw()+
  geom_smooth(method = "nls",
              formula = y ~ c*x^z,
             se = FALSE,
             method.args = list(start = list(c=111.2589, z=0.28933)))
 

```



```{r}

f.D2H_model%>%
 ggplot(aes( basal_area, height, colour = rainfall))+
  geom_point(aes(shape = as.factor(landform)))+
  geom_smooth( method = "lm")+
  scale_x_log10()+
  scale_y_log10()
```




## Basal area vs height linear mixed effect model

```{r}
#the interaction term (*) between log(basal_area) and rainfall allows the "height Vs basal areas" relationship to vary between levels of rainfall (or across sites). 

height.basal.lm <- lmer(log(height) ~ log(basal_area) * rainfall + landform + (1|plots/subplot), data = f.D2H_model, REML = T)

summary(height.basal.lm)

```






```{r}
qqnorm(residuals(height.basal.lm), pch = 1, frame = FALSE)
qqline(residuals(height.basal.lm), col = "steelblue", lwd = 2)

```






```{r}
scatter.smooth(residuals(height.basal.lm) ~ fitted(height.basal.lm))
```


```{r}
# null vs fit

height.basal.aov <- lmer(log(height) ~ log(basal_area) * rainfall + landform +  (1|plots/subplot) , data = f.D2H_model, REML = F)
height.null <- lmer(log(height) ~ log(basal_area) + landform + (1|plots/subplot), data = f.D2H_model, REML = F)

anova(height.null,height.basal.aov,test = "LRT")

```


```{r}
#differences in slope/ineraction effects

anova(height.basal.aov)


```



```{r}
#difference in intercept between high and other two sites

 pairs(emmeans(height.basal.aov, ~ rainfall))
```

```{r}

# should we keep in interaction term between rainfall and logbiomass? 

addative.basal <- lmer( log(height) ~ log(basal_area) + rainfall + landform + (1|plots/subplot), data = f.D2H_model, REML = F )

anova(addative.basal,height.basal.aov,test = "LRT")

#outputs suggest the multiplicative term is significant . BUT the AIC suggest the addative model provides a better balance between complexity and fit. 


```


```{r}
# find slope for each line 

summary(height.basal.aov, extraPar = T)


```

```{r}
vif(height.basal.aov)
```




# individual tree biomass 



```{r}

f.D2H_model %>%
  ggplot( aes (log(biomass) , fill = rainfall))+
  geom_histogram()+
  ggtitle("D2H")

f.ind_biomass %>%
  ggplot( aes(log(AGB), fill = rainfall))+
  geom_histogram()+
  ggtitle("combined")


f.paul_shrub %>%
  ggplot( aes(log(AGB), fill = rainfall))+
  geom_histogram()+
  ggtitle("shrub")

f.paul_tree%>%
  ggplot( aes (log(AGB), fill = rainfall))+
  geom_histogram()+
  ggtitle("tree")

```


```{r}

i.low.rain <- filter(D2H_model, rainfall == "low")

i.mod.rain <- filter(D2H_model, rainfall == "moderate")

i.high.rain <- filter(D2H_model, rainfall == "high")

```



```{r}
mean(i.low.rain$biomass)
mean(i.mod.rain$biomass)
mean(i.high.rain$biomass)
mean(D2H_model$biomass)
```




```{r}


ind.biomass <- f.D2H_model %>%
  ggplot(aes(rainfall, biomass, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab(bquote("biomass"(kg ~ tree^-1)))+
  xlab(" ")+
  ggtitle("d2h")+
  theme_cowplot()+
  theme(legend.position = "none",
        axis.title=element_text(size=12))


print(ind.biomass)

#ggsave("ind.biomass.png")


f.ind_biomass%>%
  ggplot(aes(rainfall, AGB, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab(bquote("biomass"(kg ~ tree^-1)))+
  xlab(" ")+
  ggtitle("combined")+
  theme_cowplot()+
  theme(legend.position = "none",
        axis.title=element_text(size=12))
  


f.pau_tree %>%
  ggplot(aes(rainfall, AGB, fill = rainfall))+
  geom_boxplot( notch = F)+
  scale_y_log10()+
  ylab(bquote("biomass"(kg ~ tree^-1)))+
  xlab(" ")+
  ggtitle("tree")+
  theme_cowplot()+
  theme(legend.position = "none",
        axis.title=element_text(size=12))

f.paul_shrub %>%
  ggplot(aes(rainfall, AGB, fill = rainfall))+
  geom_boxplot( notch = F)+
  scale_y_log10()+
  ylab(bquote("biomass"(kg ~ tree^-1)))+
  xlab(" ")+
  ggtitle("shrub")+
  theme_cowplot()+
  theme(legend.position = "none",
        axis.title=element_text(size=12))


  
```





## summed basal individaul tree anova 


```{r}

ind.tree.fit <- lmer(logAGB ~  rainfall + landform + (1|plots/subplot), data = f.ind_biomass, REML = T)

qqnorm(residuals(ind.tree.fit), pch = 1, frame = FALSE)
qqline(residuals(ind.tree.fit), col = "steelblue", lwd = 2)

 
```



```{r}

no.random <- aov(logAGB ~  rainfall + landform, data = f.ind_biomass)

qqnorm(residuals(no.random), pch = 1, frame = FALSE)
qqline(residuals(no.random), col = "steelblue", lwd = 2)

 
```
```{r}
acf(residuals(no.random))

acf(residuals(ind.tree.fit))
```



```{r}
scatter.smooth(residuals(ind.tree.fit) ~ fitted(ind.tree.fit))
```

```{r}

ind.tree.fit <- lmer(logAGB ~  rainfall * landform + (1|plots/subplot), data = f.ind_biomass, REML = F)
ind.tree.null <- lmer(logAGB ~ landform + (1|plots/subplot), data = f.ind_biomass)

anova(ind.tree.null,ind.tree.fit,test = "LRT") 

```




```{r}
pairs(emmeans(ind.tree.fit, ~ rainfall))

```

## D2H idn biomass




```{r}

sbs.tree.fit <- lmer(log_biomass ~ rainfall * landform + (1|plots/subplot), data = f.D2H_model , REML = T)

qqnorm(residuals(sbs.tree.fit), pch = 1, frame = FALSE)
qqline(residuals(sbs.tree.fit), col = "steelblue", lwd = 2)

```

```{r}
scatter.smooth(residuals(sbs.tree.fit) ~ fitted(sbs.tree.fit))
```




```{r}
acf(residuals(sbs.tree.fit))
```



```{r}

sbs.tree.fit <- lmer(log(biomass) ~ landform * rainfall + (1|plots/subplot), data = f.D2H_model , REML = F)
sbs.tree.null <- lmer(log(biomass) ~ landform + (1|plots/subplot), data = f.D2H_model)

anova(sbs.tree.null,sbs.tree.fit,test = "LRT") 

```


```{r}

pairs(emmeans(sbs.tree.fit, ~ rainfall))

```
## sensitivity analysis 

```{r}

##test robustness of structure of fixed effects 
sbs.tree.fit <- lmer(log(biomass) ~ landform * rainfall + (1|plots/subplot), data = f.D2H_model , REML = F)
sbs.tree.fit.2 <- lmer(log(biomass) ~ landform + rainfall + (1|plots/subplot), data = f.D2H_model, REML = F)

anova(sbs.tree.fit, sbs.tree.fit.2, test = "LRT")

## test robustness of structure of random effects
sbs.tree.fit.3 <- lmer(log(biomass) ~ landform * rainfall + (1|plots:subplot), data = f.D2H_model, REML = F )
anova(sbs.tree.fit, sbs.tree.fit.3, test = "LRT")



```



## summary of model perameters

```{r}
 
sbs.coef <- summary(sbs.tree.fit)$coefficients

sbs.coef
```


### parametric bootstrap 


```{r}

nBoot <- 1000
lrStat <- rep(NA, nBoot)
ft.null <- lmer(log_biomass ~ rainfall + landform + (1|plots), data =  f.D2H_model, REML = F) # null model
ft.alt <- lmer(log_biomass ~ rainfall + landform + (1|plots/subplot), data =  f.D2H_model, REML = F) # alternate model
lrObs <- 2 * logLik(ft.alt) - 2 * logLik(ft.null) # observed test stat
for (iBoot in 1:nBoot)
{
  f.D2H_model$logBiomassSim <- unlist(simulate(ft.null)) # resampled data
  bNull <- lmer(logBiomassSim ~ rainfall + landform + (1|plots), f.D2H_model, REML = F) # null model
  bAlt <- lmer(logBiomassSim ~ rainfall + landform + (1|plots/subplot), f.D2H_model, REML = F) # alternate model
  lrStat[iBoot] <- 2 * logLik(bAlt) - 2 * logLik(bNull) # resampled test stat
}
mean(lrStat > lrObs) # P-value for test of subplots effect
```


```{r}

pairs(emmeans(sbs.tree.fit, ~ rainfall))

```


# plot level biomass kg/m2


```{r}

stands_boimass <-plot_D2h %>%
  ggplot(aes( rainfall, biomass.m2, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab(bquote("biomass "(kg ~ m^-2)))+
  xlab( " d2h ")+
  ggtitle("a")+
  theme_cowplot()+
  theme(legend.position = "none")

print(stands_boimass)

plot_biomass%>%
  ggplot(aes(rainfall, biomass.m2, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab(bquote("biomass "(kg ~ m^-2)))+
  xlab( " ")+
  #ggtitle("combined")+
  theme_cowplot()+
  theme(legend.position = "none")

#ggsave("stand_biomass.png")


  p.tree_plots%>%
  ggplot(aes(rainfall, biomass.m2, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab(bquote("biomass "(kg ~ m^-2)))+
  xlab( " tree ")+
  #ggtitle("tree")+
  theme_cowplot()+
  theme(legend.position = "none")

  
  p.shrub_plots%>%
  ggplot(aes(rainfall, biomass.m2, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab(bquote("biomass "(kg ~ m^-2)))+
  xlab( " shrub ")+
  #ggtitle("tree")+
  theme_cowplot()+
  theme(legend.position = "none")

  

```





```{r}

p.low.rain <- filter(plot_D2h, rainfall == "low")

p.mod.rain <- filter(plot_D2h, rainfall == "moderate")

p.high.rain <- filter(plot_D2h, rainfall == "high")

```





```{r}

mean(p.low.rain$biomass.m2)

mean(p.mod.rain$biomass.m2)

mean(p.high.rain$biomass.m2)

mean(plot_D2h$biomass.m2)
```




## stem by stem plot biomass

```{r}

#compare main effects to interactive model 

aov.sbs.p.bio <- lmer(logbiomass.m2 ~ rainfall * landform + (1|plots), data = stand_metrics, REML = F)

null.sbs.p.bio <- lmer(logbiomass.m2 ~  landform + (1|plots), data = stand_metrics)

anova(null.sbs.p.bio, aov.sbs.p.bio, test = "LRT")

```

### model assumptions and fit

```{r}
acf(residuals(aov.sbs.p.bio))
```
```{r}
summary(aov.sbs.p.bio)
```




```{r}
qqnorm(residuals(aov.sbs.p.bio), pch = 1, frame = FALSE)
qqline(residuals(aov.sbs.p.bio), col = "steelblue", lwd = 2)

```

```{r}
scatter.smooth(residuals(aov.sbs.p.bio) ~ fitted(aov.sbs.p.bio))
```


### Compar Null to full

```{r}
aov.sbs.p.bio <- lmer(logbiomass.m2 ~ rainfall * landform + (1|plots), data = stand_metrics, REML = T)

null.sbs.p.bio <- lmer(logbiomass.m2 ~  landform + (1|plots), data = stand_metrics, REML = T)

anova(null.sbs.p.bio, aov.sbs.p.bio, test = "LRT")

pairs(emmeans(aov.sbs.p.bio, ~ rainfall))
```





## diamiter_combined paul model 


```{r}

#interactive vs main effects 

aov.p.bio <- lmer(logbiomass.m2 ~ rainfall + landform + (1|plots), data = plot_biomass, REML = F)

aov.p.int <- lmer(logbiomass.m2 ~ rainfall * landform + (1|plots), data = plot_biomass, REML = F)

anova(aov.p.int, aov.p.bio, test = "LRT")

```


### test model fit
```{r}
qqnorm(residuals(aov.p.bio), pch = 1, frame = FALSE)
qqline(residuals(aov.p.bio), col = "steelblue", lwd = 2)

```

```{r}
#questionable fit

scatter.smooth(residuals(aov.p.bio) ~ fitted(aov.p.bio))
```
### test full against null

```{r}
aov.p.bio <- lmer(logbiomass.m2 ~ rainfall * landform + (1|plots), data = plot_biomass, REML = F)

aov.p.null <- lmer(logbiomass.m2 ~ landform + (1|plots), data = plot_biomass, REML = F)

anova(aov.p.null, aov.p.bio)

pairs(emmeans(aov.p.bio, ~ rainfall))
```


## tree model

```{r}
#main effects vs interactive effects

tree.plot.main <- lmer(logbiomass.m2 ~ rainfall + landform + (1|plots), data = p.tree_plots, REML = F)

tree.plot.int <- lmer(logbiomass.m2 ~ rainfall * landform + (1|plots), data = p.tree_plots, REML = F)

anova(tree.plot.main, tree.plot.int, test = "LRT")
```


### test model fit


```{r}
qqnorm(residuals(tree.plot.int), pch = 1, frame = FALSE)
qqline(residuals(tree.plot.int), col = "steelblue", lwd = 2)

```

```{r}

scatter.smooth(residuals(tree.plot.int) ~ fitted(tree.plot.int))

```
### test full VS null

```{r}
tree.plot.fit <- lmer(logbiomass.m2 ~ rainfall * landform + (1|plots), data = p.tree_plots, REML = F)

tree.plot.null <- lmer(logbiomass.m2 ~ landform + (1|plots), data = p.tree_plots, REML = F)

anova(tree.plot.null, tree.plot.fit)

pairs(emmeans(tree.plot.fit, ~ rainfall))

```


## shrub Model 


```{r}
#main effects vs interactive effects

shrub.plot.main <- lmer(logbiomass.m2 ~ rainfall + landform + (1|plots), data = p.shrub_plots, REML = F)

shrub.plot.int <- lmer(logbiomass.m2 ~ rainfall * landform + (1|plots), data = p.shrub_plots, REML = F)

anova(shrub.plot.main, shrub.plot.int, test = "LRT")
```

### test model fit


```{r}
qqnorm(residuals(shrub.plot.int), pch = 1, frame = FALSE)
qqline(residuals(shrub.plot.int), col = "steelblue", lwd = 2)

```


```{r}

scatter.smooth(residuals(shrub.plot.int) ~ fitted(shrub.plot.int))

```


# Test full against Null

```{r}
shrub.plot.fit <- lmer(logbiomass.m2 ~ rainfall * landform + (1|plots), data = p.shrub_plots, REML = F)

shrub.plot.null <- lmer(logbiomass.m2 ~ landform + (1|plots), data = p.shrub_plots, REML = F)

anova(shrub.plot.null, shrub.plot.fit)

pairs(emmeans(shrub.plot.fit, ~ rainfall))
```


#PRODUCTIVITY 



### diam only model 


```{r}
diamiter_plot%>%
  ggplot( aes(biomass.m2, fill = rainfall ))+
  geom_histogram()
```


```{r}

diam.plot.bio <- glmmTMB(biomass.m2 ~ rainfall + landform + (1|plots) , data = p.tree_plots, family = Gamma())

diam.plot.null <- glmmTMB(biomass.m2 ~ landform + (1|plots), data = p.tree_plots, family = Gamma())

anova(diam.plot.null, diam.plot.bio, test = "LRT")

```

```{r}

pairs(emmeans(diam.plot.bio, ~ rainfall))

```

```{r}
qqnorm(residuals(diam.plot.bio), pch = 1, frame = FALSE)
qqline(residuals(diam.plot.bio), col = "steelblue", lwd = 2)
```



## average biomass VS density


```{r}
names(plot_D2h)
```
### plot average biomass vs density

```{r}
stand_metrics%>%
  ggplot( aes(subplot_density, biomass.m2, colour = rainfall))+
  geom_point(aes(shape = landform))+
  geom_smooth(method = "lm", se = F)+
  ylab(bquote("tree density "(tree/m^2)))+
  xlab(bquote("individual tree biomass " (kg/subplot)))+
  theme_cowplot()+
  theme(legend.position = "none")+
  scale_y_log10()+
  scale_x_log10()


stand_metrics%>%
  ggplot( aes(ave.biomass, subplot_density, colour = rainfall))+
  geom_point(aes(shape = landform))+
  geom_smooth(method = "lm", se = F)+
  ylab(bquote("tree density "(tree/m^2)))+
  xlab(bquote("individual tree biomass " (kg/subplot)))+
  theme_cowplot()+
  theme(legend.position = "none")+
  scale_y_log10()+
  scale_x_log10()
```


```{r}

competition <- stand_metrics%>%
  ggplot( aes((ave.biomass)^2, subplot_density, colour = rainfall))+
  geom_point(aes(shape = landform))+
  geom_smooth(method = "lm", se = F)+
  ylab(bquote("tree density "(tree/m^2)))+
  xlab(bquote("individual tree biomass " (kg/subplot)))+
  theme_cowplot()+
  theme(legend.position = "none")+
  scale_y_log10()+
  scale_x_log10()

print(competition)


  ggsave("competition.png")

```


```{r}

plot_D2h%>%
  ggplot( aes(ave.biomass, subplot_density, colour = rainfall))+
  geom_point(aes(shape = landform))
  

```



```{r}

plot_D2h%>%
  ggplot(aes(ave.biomass, colour = rainfall))+
  geom_histogram()

```



### model relationship 

```{r}

add.competition.model <- lmer(log.sub.dens ~ log(ave.biomass) + rainfall + landform +(1|plots), data = stand_metrics, REML = T)


multi.comp.model <- lmer(log.sub.dens ~ log(ave.biomass) * rainfall + landform + (1|plots), data = stand_metrics, REML = T)


anova(add.competition.model, multi.comp.model, test = "LRT")


```


```{r}

#compare fits

scatter.smooth(residuals(add.competition.model) ~ fitted(add.competition.model))

scatter.smooth(residuals(multi.comp.model) ~ fitted(multi.comp.model))
```

multiplicative model provides a better fit

```{r}
AIC(add.competition.model, multi.comp.model)
```


```{r}

# check for colinearity 

vif(add.competition.model)

```


```{r}
#look for autocorrelation

# Extract the residuals from the model
residuals <- residuals(add.competition.model)

# Perform the ACF analysis
 acf(residuals)
```


```{r}
# qqplot 

qqnorm(residuals(add.competition.model), pch = 1, frame = FALSE)
qqline(residuals(add.competition.model), col = "steelblue", lwd = 2)

```

 



```{r}
# check significance of slope 
anova(add.competition.model)
```


```{r}
summary(add.competition.model)
```




### dredging

```{r}
#rainfall 

null.rainfall.comp <- lmer(log.sub.dens ~ log(ave.biomass) + landform + (1|plots), data = stand_metrics, REML = F)

anova(null.rainfall.comp, add.competition.model)

pairs(emmeans(add.competition.model, ~ rainfall))


```

```{r}
summary(add.competition.model)
```



# stand biomass Vs density

##  plots 

```{r}
stand_metrics%>%
  ggplot( aes( biomass.m2,subplot_density, colour = rainfall))+
  geom_point(aes(shape = landform))+
  geom_smooth(method = "lm", se = F)+
 # ylab(bquote(" "(tree/m^2)))+
 # xlab(bquote("stand biomass " (kg/m2)))+
  theme_cowplot()+
  theme(legend.position = "none")+
  scale_y_log10()+
  scale_x_log10()
```


## model 

```{r}
density.biomass <- lmer(log(subplot_density) ~ rainfall * logbiomass.m2 + landform + (1|plots), data = stand_metrics)

vif(density.biomass)
```
```{r}
qqnorm(residuals(density.biomass), pch = 1, frame = FALSE)
qqline(residuals(density.biomass), col = "steelblue", lwd = 2)

```

```{r}
#compare fits

scatter.smooth(residuals(density.biomass) ~ fitted(density.biomass))

```

```{r}
#look for autocorrelation

# Extract the residuals from the model
residuals <- residuals(density.biomass)

# Perform the ACF analysis
 acf(residuals)
```

```{r}
summary(density.biomass)
```





# tree density 

```{r}

tree.density <- plot_biomass%>%
  ggplot(aes(rainfall, subplot_density, fill = rainfall))+
  geom_boxplot()+
  scale_y_log10()+
  ylab(bquote("tree density "(trees ~ m^-2)))+
  xlab(" ")+
  ggtitle("b")+
  theme_cowplot()+
  theme(legend.position = "none")


print(tree.density)

ggsave("tree.density.png")

  
```



```{r}
mean(p.low.rain$subplot_density)

mean(p.mod.rain$subplot_density)

mean(p.high.rain$subplot_density)

mean(plot_biomass$subplot_density)
```


```{r}

0.183 - 0.024

```

## lmem 

```{r}

density.fit <- lmer(log.sub.dens ~ rainfall * landform + (1|plots), data = plot_biomass)

density.null <- lmer(log.sub.dens ~  landform + (1|plots), data = plot_biomass)

anova(density.null, density.fit, test = "LRT")

pairs(emmeans(density.fit, ~ rainfall))

```


```{r}
acf(residuals(density.fit))
```



## check assumptions
```{r}
scatter.smooth(residuals(density.fit) ~ fitted(density.fit))

qqnorm(residuals(density.fit), pch = 1, frame = FALSE)
qqline(residuals(density.fit), col = "steelblue", lwd = 2)

```




# height 

```{r}

heightvsrainfall <- f.D2H_model%>%
  ggplot( aes( rainfall, height, fill = rainfall))+
  geom_boxplot(notch = F)+
  scale_y_log10()+
  ylab("height (cm)")+
  xlab(" ")+
    ggtitle("d")+
  theme_cowplot()+
   theme(axis.title=element_text(size=12),
        legend.position = "none")
  

print(heightvsrainfall)

#ggsave("heightvsrainfall.png")



```

```{r}
mean(sturt.se$height)

mean(noco.se$height)

mean(h.rain.$height)

mean(f.D2H_model$height)
```


```{r}
f.D2H_model%>%
  ggplot( aes (height, colour = rainfall))+
  geom_histogram()


```

```{r}

```
 
## tree height mixed effect model



```{r}
tree.height.lm <- lmer(log(height) ~  rainfall + landform + (1|plots/subplot), data = f.D2H_model, REML = T)
```



```{r}

qqnorm(residuals(tree.height.lm), pch = 1, frame = FALSE)
qqline(residuals(tree.height.lm), col = "steelblue", lwd = 2)


```




```{r}
scatter.smooth(residuals(tree.height.lm) ~ fitted(tree.height.lm))
```




```{r}

tree.height.fit <- lmer(log(height) ~ landform + rainfall + (1|plots/subplot), data = f.D2H_model, REML = F)
tree.height.null <- lmer(log(height) ~ landform + (1|plots/subplot), data = f.D2H_model)

anova(tree.height.null,tree.height.fit,test = "LRT")

```

```{r}
pairs(emmeans(tree.height.fit, ~ rainfall))
```

*** sig diff between low and high rainfall site. this may be caused by skewed distribution at noco***





```{r}

 sd(h.rain.$height)/sqrt(678)

 sd(noco.se$height)/sqrt(576)
 
 sd(sturt.se$height)/sqrt(517)


```





# basal area vs site

## plot basal area per site 

```{r}

basalvsrainfall <- f.d2h_model%>%
  ggplot(aes(rainfall, basal_area, colour = rainfall))+
  geom_boxplot()+
  scale_y_log10()+
  ylab("log basal area")

print(basalvsrainfall)

#ggsave("basalvsrainfall.png")



```


```{r}
mean(sturt.se$basal_area)

mean(noco.se$basal_area)

mean(h.rain.$basal_area)

mean(f.d2h_model$basal_area)
```



## basal area MELM 

```{r}

basal.site <- lmer(log(basal_area) ~ landform + rainfall + (1|plots/subplot), data = f.d2h_model, REML = T)

```



```{r}
qqnorm(residuals(basal.site), pch = 1, frame = FALSE)
qqline(residuals(basal.site), col = "steelblue", lwd = 2)
```


```{r}
scatter.smooth(residuals(basal.site) ~ fitted(basal.site))
```

```{r}

basal.site.fit <- lmer(log(basal_area) ~ landform + rainfall + (1|plots/subplot), data = f.d2h_model, REML = F)
basal.site.null <- lmer(log(basal_area) ~ landform + (1|plots/subplot), data = f.d2h_model)

anova(basal.site.null,basal.site.fit,test = "LRT")

```

```{r}
pairs(emmeans(basal.site.fit, ~ rainfall))
```

*significantly smaller BA at gundabooka compared to nocoleche. however no signifant difference between sturt and any other site. nocoleche's basal area is high due to die-back in the area disproportionately effecting smaller trees.it likely the case that differences that in basal area observed here are likely due whatever caused nocoleche's die-back rather than long term annual rainfall averages.*


# isoclines (consider statistical frontier modeling?)

```{r}

#plot level biomass
isocline <- f.D2H_model %>%
  group_by(subplot,  landform, sub_plot_area, plots , rainfall)%>%
  summarise(
    basal_area = sum(basal_area),
    subplot.n = n(),
    subplot_density = subplot.n / sub_plot_area[1],
    ave.AGB = mean(biomass)) %>% ungroup()


```

```{r}

10^seq(-3,3)

```

```{r}


contour_data <- expand_grid(
  ave.AGB = seq(1, 330, by=5),
  tot.AGB = 10^seq(-3,3)
  ) %>%
  mutate(subplot_density = tot.AGB / ave.AGB,
         rainfall=NA)

isocline.plot <- isocline %>%
ggplot(aes(ave.AGB, subplot_density, colour = rainfall))+
  geom_point(aes(shape = landform)) + 
  geom_line(data = contour_data, aes(ave.AGB, subplot_density, group = tot.AGB)) +
  scale_x_log10(limits= c(1, 330)) + scale_y_log10(limits = c(1E-3, 1E0))+
  scale_color_discrete()+
  theme_cowplot()+
  scale_color_manual(values = c("low" = "#F8766D", "moderate" = "#00BA38", "high" = "#619CFF"))+
  theme(legend.position = "none")+
  #ggtitle("c")+
  xlab( bquote("mean tree biomass " (kg)))+
  ylab( bquote( "tree density " (tree ~ m^-2)))
 

print(isocline.plot)

ggsave("isoclines.png")

```






# GRAPHS AND PANNELS


```{r}
#stand level metrics

plot_grid(stands_boimass, tree.density, competition, nrows = 3, ncol = 2, rel_widths = c(1,1))

ggsave("comparison_stand_metrics.png", width = 10, height = 10)

```



```{r}

# alometry 

plot_grid( basal.vs.heigh_plots, canopyvsheight,basalvscanopy, heightvsrainfall, ind.biomass,
          nrows = 2, ncol = 2, rel_widths = c(1,1))


ggsave("structural_comp.png", width = 10, height =10)

```



```{r}
# wood vessel analsis 

plot_grid( xylem_area, Roundness,wood.density, rel_widths = c(1,1))

ggsave("wood.Vessel.png", width = 10, height =5)

```




```{r}
# traits/organs 

plot_grid(d.massvsarea, l.thick, leaf_area, rel_widths = c(1,1))


ggsave("trait_comp.png", width = 10, height =10)

```

`

```{r}

# occurance data

plot_grid(occurence_map, rainfall_map, 
          rel_widths = c(1,1))

ggsave("distribution_comp.png", width = 10, height =5)

```





```{r}

plot_grid(l.thick, leaf_area, rel_widths = c(1,1))


ggsave("leaf_comp.png", width = 12, height =5)

```

