---
title: "R Notebook"
output: html_notebook
---

set up




```{r}

library(lme4) #mixed effect linear models

library(emmeans)

library(tidyverse)

library(readr)

library(nlme) #non linear models

library(ggpmisc)

library(cowplot)

library(multcomp)

library(MASS)

library(betareg)

library(glmmTMB)
```

```{r}
total_leaves <- read_csv("trait_data/leaves/total_leaves.csv")%>%
  mutate( rainfall = factor(rainfall, levels = c("low", "moderate", "high")))


summary(total_leaves)

```


#construct data set 

```{r}
trait_SLA <- total_leaves%>%
  mutate(
    tree_number = as.factor(tree_number),
    #mm^2/mg
    sla = leaf_area/leaf_mass,
    volume = leaf_area * leaf_thickness
    )%>%
  arrange(tree_number)%>%
  filter(!is.na(sla))%>%
  aggregate(sla~tree_number + landform + volume + leaf_area + leaf_thickness  + leaf_mass + wood_density + rainfall,mean)

summary(trait_SLA)
```


ylab("TCC on 212 "*mu~"m HDPE (mg TCC/"~m^2*") HDPE)")

labs(x = bquote('x axis'~(ring(A)^2)), y = "y axis") 

```{r}
d.massvsarea <- trait_SLA%>%
  ggplot(aes(rainfall, sla, fill = rainfall))+
  geom_boxplot()+
  scale_y_log10()+
  ylab(bquote("SLA" ~ (cm^2 ~ mg^-1)))+
  xlab("   ")+
  ggtitle("a")+ 
  theme_cowplot()+
  theme(legend.position = "none")

print(d.massvsarea)

#ggsave("d.massvsarea.png")


```

```{r}

t.rain.h <- trait_SLA%>%
  filter(rainfall == "high")

t.rain.m <- trait_SLA%>%
  filter(rainfall == "moderate")

t.rain.l <- trait_SLA%>%
  filter(rainfall == "low")


mean(t.rain.h$sla)

mean(t.rain.m$sla)

mean(t.rain.l$sla)

```


# visulise data set 

```{r}
trait_SLA%>%
  ggplot(aes(sla, colour = rainfall))+
  geom_histogram()

trait_SLA%>%
  ggplot(aes(rainfall, sla, colour = rainfall))+
  geom_boxplot()


trait_SLA%>%
  ggplot(aes(leaf_mass, leaf_area, colour = rainfall))+
  geom_point()+
  geom_smooth(method = "lm",
              se = T)+
  scale_x_log10()+
  scale_y_log10()

trait_SLA%>%
  ggplot(aes(wood_density, sla, colour = rainfall))+
  geom_point()+
  geom_smooth(method = "lm")

```




# SLA anova

```{r}
SLA.anova <- aov(log(sla) ~ rainfall + rainfall:landform, data = trait_SLA)
summary(SLA.anova)

TukeyHSD(SLA.anova)
```





## check assumptions


```{r}

qqnorm(residuals(SLA.anova), pch = 1, frame = FALSE)
qqline(residuals(SLA.anova), col = "steelblue", lwd = 2)


```


```{r}
scatter.smooth(residuals(SLA.anova) ~ fitted(SLA.anova))
```


## slope/intercept



```{r}
gunda_trait <- filter(trait_SLA, rainfall == "high")

noco_trait <- filter(trait_SLA, rainfall == "moderate")

sturt_trait <- filter(trait_SLA, rainfall == "low")

```



```{r}

high_trait <- lm(log(leaf_area) ~ log(leaf_mass), data = gunda_trait)
summary(high_trait)

exp(-1.79816)

```



```{r}

mod_trait <- lm(log(leaf_area) ~ log(leaf_mass), data = noco_trait)
summary(mod_trait)

exp(-2.68805)

```



```{r}

low_trait <- lm(log(leaf_area) ~ log(leaf_mass), data = sturt_trait)
summary(low_trait)

exp(-2.8741)

```






# leaf thickness 


```{r}

l.thick<- trait_SLA%>%
  ggplot(aes( rainfall,leaf_thickness, fill = rainfall))+
  geom_boxplot()+
  xlab("")+
  scale_y_log10()+
  ylab(bquote("thickness " (mm)))+
  ggtitle("b")+
  theme_cowplot()+
  theme(legend.position = "none")
  
print(l.thick)    

ggsave("l.thick.png")


```


```{r}

mean(trait_SLA$leaf_thickness)

mean(sturt_trait$leaf_thickness)
mean(noco_trait$leaf_thickness)
mean(gunda_trait$leaf_thickness)
 
```

```{r}

thickness.aov <- aov(log(leaf_thickness) ~  rainfall  + landform + rainfall, data = trait_SLA)

summary(thickness.aov)

TukeyHSD(thickness.aov)

```

```{r}
qqnorm(residuals(thickness.aov), pch = 1, frame = FALSE)
qqline(residuals(thickness.aov), col = "steelblue", lwd = 2)

```

```{r}
scatter.smooth(residuals(thickness.aov) ~ fitted(thickness.aov))
```






***indicates that at dryer sites leaf area to thickness relationship become negative. suggests that the biological constraints on mulga phyllode morphology are more restrictive in dryer environments leading to more homogenous phylode morphology***





# leaf area 

```{r}

leaf_area <- trait_SLA%>%
  ggplot(aes(rainfall, leaf_area, fill = rainfall))+
  scale_y_log10()+
  geom_boxplot()+
  xlab("")+
  ylab(bquote("leaf area " (cm^2)))+
  ggtitle("c")+
  theme_cowplot()+
  theme(legend.position = "none")
  
  
print(leaf_area)


```

```{r}

mean(trait_SLA$leaf_area)

mean(sturt_trait$leaf_area)
mean(noco_trait$leaf_area)
mean(gunda_trait$leaf_area)
 
```

```{r}

leaf_area.aov <- aov(log(leaf_area) ~  rainfall + rainfall + landform, data = trait_SLA)

summary(leaf_area.aov)

TukeyHSD(leaf_area.aov)

```

```{r}
qqnorm(residuals(leaf_area.aov), pch = 1, frame = FALSE)
qqline(residuals(leaf_area.aov), col = "steelblue", lwd = 2)


plot(leaf_area.aov, which = 2)

```

```{r}
scatter.smooth(residuals(leaf_area.aov) ~ fitted(leaf_area.aov))
```


# wood density



```{r}
total_wood <- read_csv("~/Thesis/r_studio/trait_data/wood/total_wood.csv")%>%
  mutate( rainfall = factor(rainfall, levels = c("low", "moderate", "high")))


summary(total_wood)

```


```{r}

wood_data <- total_wood%>%
  mutate(
    basal_area = (wood_thickness*0.5)^2*3.14159265359,
    volume = basal_area * wood_length,
    wood.dens = wood_mass/volume)

summary(wood_data)

```




```{r}

wood.density <- wood_data%>%
  ggplot(aes(rainfall, wood.dens, fill = rainfall))+
  geom_boxplot()+
  scale_y_log10()+
  ylab(bquote("wood density " (g ~ cm^-3)))+
  theme_cowplot()+
  xlab(" ")+
  ggtitle("d")+
  theme(legend.position = "none")


print(wood.density)

```



```{r}

wood.dens.aov <- aov(wood.dens ~ rainfall + landform%in%rainfall, data = wood_data)

summary(wood.dens.aov)

```







```{r}


mean(wood_data$wood.dens)

```

## check assumptions 
```{r}

qqnorm(residuals(wood.dens.aov), pch = 1, frame = FALSE)
qqline(residuals(wood.dens.aov), col = "steelblue", lwd = 2)

```



```{r}
scatter.smooth(residuals(wood.dens.aov) ~ fitted(wood.dens.aov))
```



# Wood vessel analysis

```{r}
#build data set 

vessels <- read_csv("trait_data/vessel.csv")%>%
  mutate(rainfalllow = ifelse (grepl ("STURT", name), "low", "moderate"),
          rainfallhigh = ifelse (grepl ("GUNDA", name), "high", "moderate"),
          landform = ifelse (grepl ("SHED", name), "shed", "acc"),
          rainfall = ifelse(rainfalllow == "low", "low",
                          ifelse(rainfallhigh == "high", "high", "moderate")),
          rainfall = as.factor(rainfall),
         landform = as.factor(landform),
         length = as.numeric(length),
         width = as.numeric(width),
         rainfall = factor(rainfall, levels = c("low", "moderate", "high")))%>%
  
  #extract tree as a grouping factor from the name variable
  
  mutate(
    tree = substr(name, start = 1, stop = 13))%>%
  
  #get log transformations of various variables
  mutate(log.roundness = log(roundess),
         log.area = log(area))
 
  
  f.vessels <- vessels%>% 
    filter(, l_ratio > 0.20)%>%
    filter(, roundess > 0.15)%>%
    filter(, area < 3000)%>%
    filter(, area > 60)%>%
    filter(, width > 3)%>%
     
  #turn point values into area (pixel spacial resolution 1.17um across images on average with small variance (total range = 0.014um))
  mutate(area = 1.17*area,
         length = 1.17*length,
         width = 1.17*width,
         perimeter = 1.17*perimeter)


  summary(vessels)


```




  


## vessel area  


```{r}
#diamiter
ggplot(f.vessels, aes(log.area, colour = rainfall)) +
  geom_histogram()

xylem_area <- ggplot( f.vessels, aes(rainfall, area, fill = rainfall))+
  geom_boxplot()+
  theme_cowplot()+
  scale_y_log10()+
  ylab(bquote("xylem area " (μm^2)))+
  xlab("rainfall site")+
  theme(legend.position = "none")

print(xylem_area)

#ggsave("xylem_area.png")
```

```{r}

area.model <- lmer(log.area ~ rainfall + landform + (1|tree), REML = T, data = f.vessels)


```



```{r}


qqnorm(residuals(area.model), pch = 1, frame = FALSE)
qqline(residuals(area.model), col = "steelblue", lwd = 2)
#plot(area.model, which = 2)


```

```{r}

area.model.fit <- lmer(log.area ~ rainfall + landform + (1|tree), REML = FALSE, data = f.vessels)
area.model.null <- lmer(log.area ~ landform + (1|tree), data = f.vessels)

anova(area.model.null,area.model.fit,test = "LRT") 

```


```{r}
pairs(emmeans(area.model.fit, ~ rainfall))

```

### determine effect size 

```{r}

h.vess <- filter(f.vessels, rainfall == "high")

m.vess <- filter(f.vessels, rainfall == "moderate")

l.vess <- filter(f.vessels, rainfall == "low")

mean(h.vess$area)

mean(m.vess$area)

mean(l.vess$area)

```

# roundess


```{r}
#diamiter
ggplot(f.vessels, aes( roundess, fill = rainfall)) +
  geom_histogram()

Roundness <- ggplot( f.vessels, aes(rainfall, roundess, fill = rainfall))+
  geom_boxplot()+
  theme_cowplot()+
  scale_y_log10()+
  ylab ("xylem roundness")+
  xlab("rainfall site")+
  theme(legend.position = "none")

print(Roundness)

#ggsave("Roundness.png")
```


#BETA DISTRIBUTION CONSIDER GLMM WITH CORRESPONDING LINKING FUNCTION.  

# compare different link functions 

```{r}
summary(f.vessels)
```



```{r}

gamma.glmm.roundness <- glmmTMB(roundess ~ rainfall + landform + (1|tree), data = f.vessels,family = Gamma())

beta.glmm.roundness <- glmmTMB(roundess ~ rainfall + landform + (1|tree), data = f.vessels,family = beta_family())

anova(gamma.glmm.roundness, beta.glmm.roundness, test = "LRT")

```

```{r}
possion.glmm.roundness <- glmmTMB(roundess ~ rainfall + landform + (1|tree), data = f.vessels,family = poisson())

anova(beta.glmm.roundness, possion.glmm.roundness, test = "LRT")
```



#compar to other models to check for differences in model fit

```{r}

lmem.roundness <- lmer(log.roundness ~ rainfall + landform + (1|tree), data = f.vessels)

summary(lmem.roundness)

tab_model(beta.glmm.roundness)

```


```{r}

lmem.roundness <- lmer(log.roundness ~ rainfall + landform + (1|tree), data = f.vessels)

beta.glmm.roundness <- glmmTMB(roundess ~ rainfall + landform + (1|tree), data = f.vessels,family = beta_family())

# Compare the models using likelihood ratio test (LRT)
LRT_result <- anova(lmem.roundness, beta.glmm.roundness)

# Print the results
AIC(beta.glmm.roundness, lmem.roundness)
```






```{r}

scatter.smooth(residuals(lmem.roundness) ~ fitted(lmem.roundness))


scatter.smooth(residuals(beta.glmm.roundness) ~ fitted(beta.glmm.roundness))


```


```{r}
qqnorm(residuals(lmem.roundness), pch = 1, frame = FALSE)
qqline(residuals(lmem.roundness), col = "steelblue", lwd = 2)
```


```{r}
beta_residuals <- residuals(beta.glmm.roundness)

hist(beta_residuals, breaks = 30)

```


```{r}
residuals <- residuals(beta.glmm.roundness, type = "response")

# Plot residuals
plot(beta_residuals, xlab = "Index", ylab = "Residuals", main = "Beta Residual Plot")
abline(h = 0, col = "red")
```



### test for the significace of rainfall 

```{r}

null.norm.roundess <- glmmTMB(roundess ~ landform + (1|tree), data = f.vessels, family =  beta_family())

anova(null.norm.roundess, beta.glmm.roundness, test = "LRT")

#roundness does not vary between sites

```














